### 2013-09-18 01_intro.tex
init project version

Human language communication is efficient, in that people need not always say every detail in order to be understood. Rather, speakers only say what is most relevant, and listeners can seemingly effortlessly grasp the intended meaning beyond what is literally said. This ability to do pragmatic inference has been long studied and the \emph{conversational implicature} theory by Grice \cite{Grice} is one of the most prominent theories in the field. However, since it is hard to formalize the \emph{Cooperative Principle} and the \emph{Conversational Maxims}, Grice's theory is not precise enough to give empirically testable predictions, especially quantitative predictions.

The Bayesian \emph{Rational Speech Act} (RSA) model attempts to address this issue by using information-theoretic notions together with the Bayesian inference framework. It has been shown that the model could yield quantitative predictions that are highly correlated to actual human judgments \cite{Frank}.

Despite the RSA model's theoretical advantage in that it provides us with quantitative predictions and its empirical success, if we analyze the design of the model, we will find several choices that are not \emph{prima facie} obvious and thus need further clarification. These choices have their alternatives in the closely related game-theoretic approaches to pragmatics (e.g. \cite{Benz2007,Jager2013}), which have similar but slightly different conceptual motivations. Hence it is important to systematically compare the original RSA model with all these alternatives and test their predictions on empirical data, to gain a better understanding of the nature of human pragmatic inference.

The outline of the paper is as follows. We first introduce the simple referential communication game investigated in \cite{Frank} and the original RSA model proposed there. We then analyze a few design choices in this model and introduce their alternatives in game-theoretic pragmatics. Next we report the result of an experiment similar to the original one, and compare the predictions of different models to the new data. Finally we report on an additional experiment that suggests the sparsity of higher-order inference in actual human pragmatic reasoning.

### 2013-09-18 02_RSA.tex
init project version

Figure \ref{context} is an example of the referential game used in studying human pragmatic inference \cite{Frank}. In a referential game there is a set of objects called the \emph{context}. The speaker in the game is asked to refer to one of the objects (the \emph{target}) by uttering an expression (typically a word) to the listener. The listener, who has no idea which object is the target, needs to recover the speaker's intended referent based on the speaker's utterance.

\begin{figure}[h] 
  \centering
  \subfigure[The context]{\label{context}
  \begin{tikzpicture}
     \path (-1.8, 0) node [shape=rectangle, draw=green!65 ,fill= green!65 ,minimum size=28]{}
           ( 0, 0) node  [shape=circle, draw=green!65, fill= green!65,minimum size=30] {}
           ( 1.8, 0) node [shape=circle, draw=blue!65, fill=blue!65  ,minimum size=30] {}
           ( -1.8, -0.8) node {$t_{1}$}
           ( -0.0, -0.8) node {$t_{2}$}
           (  1.8, -0.8) node {$t_{3}$}
     ;

  \end{tikzpicture}
  }  
  \hspace{5mm}
  \subfigure[The vocabulary and the truth table]{\label{VaTT}
  \begin{tikzpicture}
     \path (0,0) node[] 
     {\begin{tabular}{r|ccc}
                         & \quad $t_{1}$\ \  &\quad $t_{2}$\ \  &\quad $t_{3}$ \ \   \\
    \hline
   ``Green'':  $m_{1}$ \quad  &  $1$        &   $1$       & $0$      \\
   ``Circle'': $m_{2}$  \quad &  $0$        &   $1$       & $1$      \\ 
   ``Square'': $m_3$ \quad    &  $1$      &   $0$       & $0$      \\
   ``Blue'':   $m_4$  \quad  &  $0$        &   $0$       & $1$     
     \end{tabular}
     } ;
  \end{tikzpicture}
  }
  \caption{A simple referential game}
\end{figure}

Let us first try to play this game ourselves, to get a better sense of what it is about and why it is relevant to human pragmatic reasoning. For simplicity let us assume that the utterances that the speaker is allowed to use are from a \emph{vocabulary} which is commonly known between the speaker and the listener. The content of the vocabulary and the (truth-conditional/literal) meaning of each word are shown in Figure \ref{VaTT}. Suppose we play as the speaker and the target is $t_1$, the green square. We can either use ``Square'' or ``Green'' to literally truthfully refer to it, but it seems more prudent to use ``Square'', as there is only one square in the context and thus the listener can easily identify it, whereas there are two green objects, which makes ``Green'' ambiguous. In terms of the Gricean \emph{Maxim of Quantity}, we should use ``Square'' because it is as maximally informative as is required. Similarly, we should use ``Blue'' to refer to $t_3$, the blue circle. However,  using ``Blue'' to refer to the blue circle might intuitively sound a little unnatural, as color terms are usually used as adjectives\footnote{They are used as nouns mostly to refer to the colors themselves.}, while we usually use nouns to refer to concrete objects. This inclination becomes more evident when we want to refer to $t_2$, the green circle. While ``Green'' and ``Circle'' are equally ambiguous in this case, we might nevertheless prefer the latter.

Now let us turn to play as the listener. If we hear ``Square'' or ``Blue'', we easily know the intended referent as there is no ambiguity, but what if we hear ``Circle'' (or ``Green'')? There are two circles in the context that we need to choose from. On the one hand, the blue circle, having the unique color in the context, seems to be perceptually dominant and thus easily captures our attention. On the other hand, from the previous analysis we know that if the blue circle were the intended referent, the speaker could have chosen ``Blue'' which is not ambiguous and thus more informative. Hence the listener needs to balance two sources of information, i.e. the (presumably subconscious) perceptual salience of different objects and the rational expectation of the likelihood of the speaker making the utterance for each object. The latter line of reasoning is crucial to the classic Gricean account of scalar implicatures, and the major challenges in terms of formal modeling are how to quantify  notions such as informativeness and how different pieces of information should be integrated.

The RSA model addresses the problems by using information-theoretic concepts to measure informativeness and Bayesian inference to integrate different sources of information \cite{Frank}.

In order to measure the informativeness of an utterance, the RSA model starts with the literal listener who upon receiving an utterance $m$ does a conditioning on its literal meaning:
\begin{equation}\label{Bayesian_literal}
\rho_0(t \mid m) = \mathcal{U}(t\mid \intp{m}) ,
\end{equation}
where $\mathcal{U}$ is a uniform distribution over all the possible referents. Now the informativeness of utterance $m$ for the intended referent $t$ can be measured as the negative Kullback-Leibler divergence of the induced literal listener's belief $\rho_0$ from the speaker's own belief $\delta_t$:
\begin{equation} \label{Bayesian_sender_informativeness}
\mbox{Info}(m,t)=-\mbox{KL}(\delta_t \| \rho_0)=-\sum_{t'}\delta_t(t')\log\left(\frac{\delta_t(t')}{\rho_0(t')}\right),
\end{equation}
where $\delta_t$ is a delta distribution with all the probability mass on $t$, as the speaker knows her intended referent:
\begin{equation} \label{Bayesian_sender_delta}
\delta_t(t')=\left\{ \begin{array}{ll}
1 & \mbox{if } t' = t \\
0 & \mbox{otherwise}
\end{array}\right. \enspace .
\end{equation}

The speaker acts sub-optimally by choosing the utterance that soft-maximizes her expected utility, which is defined as the informativeness of the utterance subtracted by its cost:
\begin{equation} \label{Bayesian_sender_softmax}
\sigma(m \mid t) \propto \exp(\lambda_\mathrm{S}\cdot \mbox{U}(m,t))=\exp(\lambda_\mathrm{S} \cdot (\mbox{Info}(m,t)-\mbox{Cost}(m))),
\end{equation}
where $\lambda_\mathrm{S}$ is a parameter measuring the speaker's degree of rationality, i.e. to what extent the speaker sticks to the strict optimum. The cost term is used to encode preference in different utterances, be it about the utterances' lengths or syntactic categories.

From (\ref{Bayesian_literal})-(\ref{Bayesian_sender_softmax}) we obtain the speaker's production rule:

\begin{equation} \label{Bayesian_speaker}
\sigma(m \mid t) \propto \exp(\lambda_\mathrm{S} \cdot (\log \mathcal{U}(t\mid \intp{m})-\mbox{Cost}(m))) \enspace .
\end{equation}

The pragmatic listener in the RSA model, upon receiving the utterance $m$, does a Bayes update on his prior belief $\mathcal{S}(t)$ by using his knowledge of the speaker's behavior (\ref{Bayesian_speaker}):
\begin{equation} \label{Bayesian_rec_update}
\rho(t \mid m) \propto \mathcal{S}(t)\cdot \sigma(m \mid t) \enspace .  
\end{equation}
The Bayes rule helps naturally integrate the perceptual salience of each object, which is treated as the prior $\mathcal{S}(t)$ and can be empirically measured, with listener's expectation of the speaker being informative, which is incorporated as the likelihood, thus addressing the previously mentioned challenge of balancing different sources of information. Setting $\lambda=1$ and $\mbox{Cost}(m)=0$ for all $m$, \cite{Frank} obtained a highly significant correlation between the model prediction and the experiment data.

### 2013-09-18 03_alternatives.tex
init project version

The RSA model has the theoretical advantage in that it is quantitative, and it has been shown to be able to fit the empirical data well. However, if we want to gain more insight about communication involving referential expressions from its success, it will be worthwhile to examine the model carefully to pin down the major components, spell out the design choices and the underlying assumptions, test their contribution to the predictive power of the model through comparison with the alternatives and scrutinize their implications thereof. In this section, we will carry out a conceptual analysis of the RSA model by comparing it to the closely related game-theoretic approaches to pragmatics, to define a class of alternative models. We will then compare the predictions of these models to the empirical data in later sections.

First and foremost, there are three closely related notions, i.e. \emph{belief}, \emph{goal} and \emph{action}, that play a crucial role in the RSA as well as game-theoretic models. The relation and distinction between them can be best illustrated by looking at (\ref{Bayesian_speaker}):
$$\sigma(m \mid t) \propto \exp(\lambda_\mathrm{S} \cdot (\log \mathcal{U}(t\mid \intp{m})-\mbox{Cost}(m))) \enspace .$$
The term $\mathcal{U}(t\mid \intp{m})$ is the speaker's \emph{belief} about the listener's interpretation of the utterance. The expected utility $\mbox{U}(m,t)=\log \mathcal{U}(t\mid \intp{m})-\mbox{Cost}(m)$ describes the speaker's \emph{goal} of communication, i.e. inducing a belief that is as close as his own with minimal effort. The production rule $\sigma(m \mid t)$ on the left-hand side specifies the speaker's \emph{action}, as it defines the probability the speaker actually chooses an action, i.e. uttering a particular expression. The soft-max rule connects these notions, as a (sub-)rational agent's action depends on his belief and goal. While the RSA and game-theoretic models all share this general architecture in their designs, they vary in the specific interpretations of the three ingredients, which reflect different views and emphasis on language and communication and give rise to a series of interesting and important empirical questions. Again, let us continue with the speaker model (\ref{Bayesian_speaker}) to illustrate some points of divergence and formally define the corresponding alternatives.

First of all, although \cite{Frank} claims that the empirically measured perceptual salience is used to capture the notion of common knowledge between the speaker and the listener, it turns out that the speaker's production rule (\ref{Bayesian_speaker}) simply does not take the listener's salience prior $\mathcal{S}(t)$ into account in her literal listener model (\ref{Bayesian_literal}). This makes it unclear in what sense the speaker knows the listener's perceptual salience. A natural variant, which is used in some game-theoretic models, would be to replace the literal listener's uniform distribution $\mathcal{U}(t)$ in (\ref{Bayesian_literal}) with the salience prior $\mathcal{S}(t)$. This will lead to the alternative production rule:
\begin{equation} \label{Bayesian_speaker_salience}
\sigma_\mathcal{S}(m \mid t) \propto \exp(\lambda_\mathrm{S} \cdot (\log \mathcal{S}(t\mid \intp{m})-\mbox{Cost}(m))) \enspace .
\end{equation}

Secondly, the informativeness of an utterance $m$, which is a crucial part of the communication goal, is measured in the RSA model in terms of how close the induced belief of the literal listener is from that of the speaker's own belief. This means that the RSA model sees the goal of communication as conveying belief. While it is normally true that language does convey the belief of the speaker, it is questionable at least in this referential scenario whether letting the listener form a proper belief is the ultimate goal of the communication. After all, if the speaker wants to refer to something, it seems that in the end what matters is whether the listener actually picks out the intended referent successfully (e.g., when the speaker wants the listener to pass something). We might call this view \emph{action-oriented}, in contrast to the \emph{belief-oriented} view of communication which the RSA model adopts, as it interprets the goal of communication as invoking the intended action rather than forming an accurate belief. \footnote{Note that the two views are definitely intimately related, despite our emphasis on their differences here. An accurate belief is usually helpful to choosing a desired action. In fact, forming a belief can itself be seen as an action that the listener performs, which might well be all that the speaker cares about in many everyday situations, e.g. casual talks or responses to inquiries of information. Also, the relevant actions might vary greatly across situations which makes belief formation a practical alternative to serve as the goal. Nevertheless, they are still essentially different views in spite of the fact that they often coincide in practice. The referential communication scenario we investigate in this paper reflects the distinction.} Thus, according to this view, the informativeness of a message would be measured as the probability of the listener choosing the intended referent. Formally speaking, we have the action-oriented speaker model

\begin{equation} \label{Bayesian_speaker_action}
\sigma_\mathrm{a}(m \mid t) \propto \exp(\lambda_\mathrm{S} \cdot (\rho_0(t\mid m)-\mbox{Cost}(m))) \enspace .
\end{equation}

Hence we have four types of speaker models that differ in either the speaker's belief about the literal listener, or the speaker's goal of communication. We now introduce a uniform notation $\sigma_{xy},\ x\in\{a,b\},y\in\{\mathcal{U},\mathcal{S}\}$ for them:
\begin{equation} \label{Bayesian_speaker_uniform_action}
\sigma_{\mathrm{a}y}(m \mid t) \propto \exp(\lambda_\mathrm{S} \cdot ( y(t\mid m)-\mbox{Cost}(m))),
\end{equation}
\begin{equation} \label{Bayesian_speaker_uniform_belief}
\sigma_{\mathrm{b}y}(m \mid t) \propto \exp(\lambda_\mathrm{S} \cdot (\log y(t\mid m)-\mbox{Cost}(m))),
\end{equation}
where $\mathcal{U}$ is the uniform prior and $\mathcal{S}$ is the salience prior. For example, in the original RSA model, the speaker does not take listener's salience prior into account and he has a belief-oriented goal of communication, thus it will be denoted as $\sigma_{\mathrm{b}\mathcal{U}}$

Finally, the original RSA speaker model in \cite{Frank} has $\mbox{Cost}(m)=0$ for all $m$, which means that the potential speaker preference in different utterances is not taken into account. As we previously pointed out, our intuition seems to suggest a preference for nouns over adjectives as referential expressions. Since it is an empirical question whether, to what direction or to what extent such a preference exists, we leave open all the possibilities. Technically speaking, since only the difference in costs matters, we define the cost function using a constant $c\in \mathbb{R}$
\begin{equation} \label{cost}
\mbox{Cost}(m)=c \mbox{ if } m \mbox{ is an adjective and }0 \mbox{ otherwise} \enspace . 
\end{equation}
If $c>0$ it means there is a preference for nouns and if $c<0$ then the preference is for adjectives. No preference exists if $c=0$.

Now we turn to the listener model (\ref{Bayesian_rec_update}), where the boundary between belief, goal and action becomes less clear:
$$\rho(t \mid m) \propto \mathcal{S}(t)\cdot \sigma(m \mid t) \enspace . $$

Let us start from the relatively clear part. As pointed out previously, the likelihood term $\sigma(m \mid t)$ is the listener's belief about how the speaker behaves. One natural question is to what extent the listener's belief correlate with the actual production. We thus treat the speaker's production term $\sigma(m \mid t)$ in the listener's model as a parameter, making the association with a belief about production of the listener model explicit:
\begin{equation} \label{listener-production}
\rho(\sigma_{xy})(t \mid m) \propto \mathcal{S}(t)\cdot \sigma_{xy}(m \mid t) \enspace .
\end{equation}
Note that the speaker's production rule has two parameters $\lambda_\mathrm{S},c$ which are also be included in the above specification of the listener model.

Next, even though our intuition suggests different objects have different perceptual salience and thus might affect our judgment, it is after all an empirical question whether it is relevant in the interpretation of referential expressions. It is in principle possible that the listener does not take into account the perceptual salience in his reasoning, which means he has a uniform prior over the referents:
\begin{equation} \label{listener-production-uniform}
\rho_{\mathcal{U}}(\sigma_{xy})(t \mid m) \propto \mathcal{U}(t)\cdot \sigma_{xy}(m \mid t) \enspace .
\end{equation}

Finally, since the RSA model adopts a belief-oriented view on the goal of communication, its listener model only consists of the belief about the intended referent after hearing the utterance, obtained by a Bayes update. However, according to the action-oriented view, this is not quite the end of the story. The listener often needs to decide the exact referent of a referential expression. (Again, consider the case in which the listener is asked to pass something.) Hence in such cases the listener will choose an object that he (soft-)maximally believes to be the intended referent. Formally speaking, the action-oriented listener model would be
\begin{equation} \label{listener-action}
\rho_{\mathrm{a}v}(\sigma_{xy})(t \mid m) \propto \exp(\lambda_\mathrm{L} \cdot \rho_{\mathrm{b}v}(\sigma_{xy})(t \mid m) ),
\end{equation}
where $v\in\{\mathcal{U},\mathcal{S}\}$, $\lambda_\mathrm{L}$ is the parameter measuring the listener's degree of rationality, and $\rho_{\mathrm{b}v}$ is the belief-oriented model that does the Bayes update:
\begin{equation} \label{listener-belief}
\rho_{\mathrm{b}v}(\sigma_{xy})(t \mid m) \propto v(t)\cdot \sigma_{xy}(m \mid t) \enspace .
\end{equation}
For instance, the original RSA listener model is a belief-oriented one with the perceptual salience as prior, whose belief about the speaker is $\sigma_{\mathrm{b}\mathcal{U}}$, hence it is denoted as $\rho_{\mathrm{b}\mathcal{S}}(\sigma_{\mathrm{b}\mathcal{U}})$.

### 2013-09-18 04_experiment.tex
init project version

In the previous section we analyzed the design of the original RSA model, spelled out the relation and distinction between belief, goal and action, and proposed a familiy of alternative models based on different interpretation of these notions. This gives rise to a series of interesting empirical questions regarding the underlying assumptions of the models. In order to gain insight into these questions by comparing the predictive power of different models, we conducted the following experiment to collect empirical data, which we will use to test the predictions of different models in the next section.

There are four major designs in our own experiment that differ from the original experiment in \cite{Frank}. First of all, the original experiment used the betting paradigm, i.e. participants were asked to bet over a range of options and they were instructed that the amount of money that is bet on an option should reflect their confidence that the option is correct. Even though the betting paradigm has the merit of providing us with graded reponses from each individual participant, the caveat is that it is unclear whether it measures the belief or the action. This can lead to confusion when we are to fit the model predictions to the empirical data without knowing whether they are directly comparable. In addition, since the betting paradigm is more or less introspective in nature, it tends to be not very accurate. Thus we used forced choice in our experiment instead, which clearly measures the action and does not rely on introspection.
Secondly, since we decided to investigate the influence of the speaker's preference as well as the listener's perceptual salience, we focused on contexts equivalent to Figure \ref{context}, which are most typical in Gricean accounts of scalar implicature, and examined different features separately.
Thirdly, we only included the features of color and shape to ensure the vocabulary to be common knowledge, excluding the feature of texture. Finally, we did not use dotted line surrounding an object as the way to indicate the target to the speaker, as it could emphasize the feature of shape and thus add confound in the experiment. Instead we used a small arrow pointing to the target. Despite these differences, the remaining aspects in our experiment were almost identical to those of the original one, e.g. the phrasing of the instructions. We also counterbalanced the positions,  colors, and word orders in each condition.

The results of the three parts of our experiment are as follows.

\subsection{Prior Estimation}
We recruited 240 US participants via Amazon Mechanical Turk and presented them with a context of three objects (Figure \ref{exp_context} as an example).

\begin{figure}[htb] 
  \centering
  \subfigure[A sample context]{\label{exp_context}
  \begin{tikzpicture}
     \path (-1.4, 0) node [shape=rectangle, draw=green!65 ,fill= green!65 ,minimum size=28]{}
           ( 0, 0) node  [shape=circle, draw=green!65, fill= green!65,minimum size=30] {}
           ( 1.4, 0) node [shape=circle, draw=blue!65, fill=blue!65  ,minimum size=30] {}
     ;

  \end{tikzpicture}

  }  
  \hspace{5mm}
  \subfigure[Choice distribution]{\label{table:prior}
  \begin{tikzpicture}
     \path (0,0) node[] 
     {\begin{tabular}{c|ccc}
    \quad &\textcolor{green!65}{\Large{$\blacksquare$}}\quad & \textcolor{green!65}{\Huge{$\bullet$}}\quad & \textcolor{blue!65}{\Huge{$\bullet$}} \\ 
     \hline
     & $71$        &   $30$       & $139$        
     \end{tabular}
     } ;
  \end{tikzpicture}
  }
  \caption{Prior Estimation}
\end{figure}

Participants were instructed to imagine that someone was talking to them and referring to one of the objects, using a foreign language they do not understand. Then they were asked to choose the object they thought that person was talking about. The numbers of participants choosing each object are shown in Figure \ref{table:prior}. We will use this empirical distribution to estimate the perceptual salience prior of each object.

\subsection{Speaker Conditions}

We recruited 432 US participants via Amazon Mechanical Turk in the speaker conditions, 144 in each condition. The participants were presented with a context similar to Figure \ref{exp_context}, with a small arrow indicating the target. The participants were then asked to imagine they were talking to someone else and to choose from the two literally true words to refer to the target object.

\begin{table}[htb]   
  \caption{Speaker Conditions}
  \centering
  \begin{tabular}{c|cccc}
   Target  & ``Green'' & ``Square'' & ``Blue'' & ``Circle'' \\ 
     \hline
\textcolor{green!65}{\Large{$\blacksquare$}}    & $9$        &   $135$   & -- & -- \\
\textcolor{blue!65}{\Huge{$\bullet$}}           & --        &   --      & $119$ & $25$\\
\textcolor{green!65}{\Huge{$\bullet$}}          & $63$        &   --    &  --   & $81$             
  \end{tabular}

  \label{table:speaker}
\end{table}

The numbers of participants choosing each message in each condition are shown in Table \ref{table:speaker}. Note that when the target was the object with the unique shape (as in the first row of the table), the feature of shape (``Square'' in the example) should be the optimal utterance because the listener could uniquely identify it. Similarly when the target was the object with the unique color (the second row), the optimal utterance would be the feature of color (``Blue''). When the target was the object with both features shared, both features should be equally ambiguous because of the context's symmetric nature.

From the data in Table \ref{table:speaker}, we can see that the speakers tend to choose the optimal feature more often when the target has the unique shape than when it has the unique color ($\chi^2=8.5, p<.01$). Even though they seem to prefer the feature of shape when the target's both features are shared, the difference is not statistically significant from uniform random choice ($\chi^2=2.25, p=.13$).

\subsection{Listener Conditions}

We recruited 360 US participants via Amazon Mechanical Turk in the speaker conditions, 180 in each condition. The participants were presented with a context similar to Figure \ref{exp_context}, together with a word denoting either of the shared features (color or shape). The participants were asked to imagine someone else was talking to them and he used the word to refer to one of the objects. Then they were asked to choose the object they thought that person was talking about.

\begin{table}[htb] 
  \caption{Listener Conditions}
  \centering
\begin{tabular}{c|ccc}
   & \quad \textcolor{green!65}{\Large{$\blacksquare$}}&  \textcolor{green!65}{\Huge{$\bullet$}}& \textcolor{blue!65}{\Huge{$\bullet$}} \\ 
     \hline
 ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & $0$      \\
 ``Circle''   &  \quad $1$        &    $62$        & $117$
  \end{tabular}

  \label{table:listener}
\end{table}

The numbers of participants choosing each object upon hearing each word are shown in Table \ref{table:listener}. Note that in either case, the object with both features shared (the green circle) is what the Gricean pragmatic account predicts to be the target.

The result in Table \ref{table:listener} shows that when the message is ``Green'', listeners prefer the green circle which is predicted by the Gricean pragmatics, while they tend to stick to the blue circle which is more perceptually salient when the message is ``Circle''\footnote{Admittedly it is strange that one participant chose the square when the message was ``Circle''. However, since he passed the attention check we just include his data as it is.}. The behavioral patterns in both conditions are significantly different from uniform random choice, and they significantly differ from each other in respect of whether they conform to the predictions by Gricean pragmatic theory.

### 2013-09-18 06_conclusion.tex
init project version

Our main contributions are as follows: (1) We emphasized the distinction between belief, goal and action, systematically examined the RSA model, connected it to the game-theoretic approaches. (2) We tested the RSA model with a family of alternatives on the forced choice experiment and drew attention to the speaker's preference. Our data seems to suggest a different story about common knowledge --- what is commonly known is the Gricean pragmatic components, while both sides have his own private bias unknown to the other. (3) We explored the possibility of speakers doing second-order pragmatic inference and showed that it seems hard for them to do so.

Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to further improvement in predictive power. Of course, more data from a wide range of further experiments are needed for a full understanding of pragmatic inference about referential expressions.

### 2013-09-18 main.tex
init project version

\documentclass{llncs}

% ======= Packages ======

%    ### General ###
%\usepackage{latexsym}
\usepackage{enumerate}

\usepackage{amssymb,amsmath}   %
%\usepackage{amsthm} % theorem environment

%    ### Format ###
%\usepackage{a4wide}     %% wide page
%\usepackage{times}      %% small fonts
%\usepackage{fullpage}

%    ### Logic and Theoretical Computer Science ###
\usepackage{stmaryrd}
\usepackage[all, knot]{xy}%%xy-pics
\usepackage{verbatim}
\usepackage[english]{babel}
\usepackage{graphicx}

%    ### Game Theory ###

%\usepackage{xspace}
%\usepackage{tikz}                %avoid multiple call !!
  %\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}

%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
%\setlength\PreviewBorder{5pt}

%\usepackage{lgcommands}
\usepackage{tikz,subfigure}
   \usetikzlibrary{shapes.geometric}

%    ### Linguistics ###
%\usepackage{tipa}   % International Phonetic Alphabet

%\usepackage{qtree}              %basicly   qtree and tikz-qtree do the samething
%\usepackage{tikz}                  % avoid multiple call !!
%\usepackage{tikz-qtree}

%\usepackage{gb4e}         % NOTE!!!   This should be the LAST  usepackage   call!!!!!!!
%\usepackage{lingmacros}
% ======== Definitions ========

\newcommand{\intp}[1]{\llbracket #1 \rrbracket}         %interpretation

% ======== Title ========

\title{Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning}
\author{Michael Franke \and Ciyang Qing}
\institute{Institute for Logic, Language and Computation}

% ======== Content ========

\begin{document}

\maketitle

\begin{abstract}
Recent developments in Bayesian experimental pragmatics have received much attention, as it provides us with a quantitative model formalizing core concepts in traditional pragmatic theories and the predictions it makes fit empirical data nicely. Despite its theoretical advantage and empirical success, there are still several conceptual issues in the Bayesian Rational Speech Act (RSA) model that need further clarification. In this paper, we analyze a few design choices made in RSA that have intuitively equally plausible alternatives motivated from related game-theoretic approaches to pragmatics, and compare the original RSA model with these alternatives to empirical data. The result suggests that while RSA is better than most other variations, there is still room for improvement.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\input{01_intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Referential Games and the RSA model}

\input{02_RSA}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternatives in Design Choices}

\input{03_alternatives}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Experiment}

\input{04_experiment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Comparison}

\input{05_model_comparison}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

\input{06_conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{splncs}
\bibliography{variations_ref}

\end{document}

### 2013-09-24 04_experiment.tex

@@
 In order to gain insight into these questions by comparing the predictive power of different models, we conducted the following experiment to collect empirical data, which we will use to test the predictions of different models in the next section.
+
+\subsection*{Participants}
+
+We recruited 1032 US Participants via Amazon Mechanical Turk, an online crowd-sourcing tool.
+Each participant played a one-shot referential game either as a speaker or as a listener and received a payment of 5 cents.
+If a participant contributed multiple trials, only the first was included.
+All participants passed the attention check described below.
+
+\subsection*{Materials and Procedure}
+
+Each participant saw a context consisting of three objects (Figure \ref{exp_context} as an example) and reported on the number of objects in the context as an attention check.
+Participants in the listener conditions were told to imagine someone was talking to them and used some utterance (depending on the exact condition) to refer to exactly one of the objects in the context.
+Then they were asked to choose what they thought to be the intended referent.
+Participants in the speaker conditions were told to imagine they were talking to someone and they wanted to refer to one of the objects (indicated by a small red arrow\footnote{Participants were told that the arrow was only for illustration and thus the person they were talking to could not see it.}).
+Then they were asked to choose between two words to refer to it.
+
+\begin{figure}[htb] 
+  \centering
+  \begin{tikzpicture}
+     \path (-1.4, 0) node [shape=rectangle, draw=green!65 ,fill= green!65 ,minimum size=28]{}
+           ( 0, 0) node  [shape=circle, draw=green!65, fill= green!65,minimum size=30] {}
+           ( 1.4, 0) node [shape=circle, draw=blue!65, fill=blue!65  ,minimum size=30] {}
+     ;
+
+  \end{tikzpicture}
+
+  \caption{A sample context}\label{exp_context}
+\end{figure}
+
+Each context had a square and a circle sharing the same color, and another circle having a different color.
+Hence each object had two features: shape (square/circle) and color (green/blue) and each context had two ambiguous (shared by two objects) features.
+In the listener salience condition, participants were told that the person talking to them was using a foreign language they could not understand, and in the other two listener conditions, the utterances used were the ambiguous feature of shape and color, respectively.
+In the speaker conditions, the target could be any of the three objects: the one with the unique color, the one with the unique shape and the one with both features shared.
+The two words for the participants to choose between were the features of color and shape of the target object.
+
+The positions and colors of the objects, as well as the orders of the candidate words in the speaker conditions, were counterbalanced.
+
+\subsection*{Design}
 
@@
 Instead we used a small arrow pointing to the target.
-Despite these differences, the remaining aspects in our experiment were almost identical to those of the original one, e.g.
-the phrasing of the instructions.
-We also counterbalanced the positions,  colors, and word orders in each condition.
 
-The results of the three parts of our experiment are as follows.
+Despite these differences, the remaining aspects in our experiment were almost identical to those of the original one, e.g. the phrasing of the instructions.
 
-\subsection{Prior Estimation}
-We recruited 240 US participants via Amazon Mechanical Turk and presented them with a context of three objects (Figure \ref{exp_context} as an example).
+\
 
-\begin{figure}[htb] 
-  \centering
-  \subfigure[A sample context]{\label{exp_context}
-  \begin{tikzpicture}
-     \path (-1.4, 0) node [shape=rectangle, draw=green!65 ,fill= green!65 ,minimum size=28]{}
-           ( 0, 0) node  [shape=circle, draw=green!65, fill= green!65,minimum size=30] {}
-           ( 1.4, 0) node [shape=circle, draw=blue!65, fill=blue!65  ,minimum size=30] {}
-     ;
-
-  \end{tikzpicture}
-
-  }  
-  \hspace{5mm}
-  \subfigure[Choice distribution]{\label{table:prior}
-  \begin{tikzpicture}
-     \path (0,0) node[] 
-     {\begin{tabular}{c|ccc}
-    \quad &\textcolor{green!65}{\Large{$\blacksquare$}}\quad & \textcolor{green!65}{\Huge{$\bullet$}}\quad & \textcolor{blue!65}{\Huge{$\bullet$}} \\ 
-     \hline
-     & $71$        &   $30$       & $139$        
-     \end{tabular}
-     } ;
-  \end{tikzpicture}
-  }
-  \caption{Prior Estimation}
-\end{figure}
-
-Participants were instructed to imagine that someone was talking to them and referring to one of the objects, using a foreign language they do not understand.
-Then they were asked to choose the object they thought that person was talking about.
-The numbers of participants choosing each object are shown in Figure \ref{table:prior}.
-We will use this empirical distribution to estimate the perceptual salience prior of each object.
+The results of the speaker and listener conditions of our experiment are shown as follows.
 
@@
 
-We recruited 432 US participants via Amazon Mechanical Turk in the speaker conditions, 144 in each condition.
-The participants were presented with a context similar to Figure \ref{exp_context}, with a small arrow indicating the target.
-The participants were then asked to imagine they were talking to someone else and to choose from the two literally true words to refer to the target object.
+There were 432 participants in the speaker conditions, 144 in each condition.
+The numbers of participants choosing each word in each condition are shown in Table \ref{table:speaker}.
+Note that when the target was the object with the unique shape (as in the first row of the table), the feature of shape (``Square'' in the example) should be the optimal utterance because the listener could uniquely identify it.
+Similarly when the target was the object with the unique color (the second row), the optimal utterance would be the feature of color (``Blue'').
+When the target was the object with both features shared, both features should be equally ambiguous because of the context's symmetric nature.
 
@@
   \centering
-  \begin{tabular}{c|cccc}
-   Target  & ``Green'' & ``Square'' & ``Blue'' & ``Circle'' \\ 
+  \begin{tabular}{c|ccccc}
+   Target  & ``Green'' & ``Square'' & ``Blue'' & ``Circle'' & Total\\ 
      \hline
-\textcolor{green!65}{\Large{$\blacksquare$}}    & $9$        &   $135$   & -- & -- \\
-\textcolor{blue!65}{\Huge{$\bullet$}}           & --        &   --      & $119$ & $25$\\
-\textcolor{green!65}{\Huge{$\bullet$}}          & $63$        &   --    &  --   & $81$             
+\textcolor{green!65}{\Large{$\blacksquare$}}    & $9$        &   $135$   & -- & -- &144\\
+\textcolor{blue!65}{\Huge{$\bullet$}}           & --        &   --      & $119$ & $25$ & 144\\
+\textcolor{green!65}{\Huge{$\bullet$}}          & $63$        &   --    &  --   & $81$             &144
   \end{tabular}
@@
 \end{table}
-
-The numbers of participants choosing each message in each condition are shown in Table \ref{table:speaker}.
-Note that when the target was the object with the unique shape (as in the first row of the table), the feature of shape (``Square'' in the example) should be the optimal utterance because the listener could uniquely identify it.
-Similarly when the target was the object with the unique color (the second row), the optimal utterance would be the feature of color (``Blue'').
-When the target was the object with both features shared, both features should be equally ambiguous because of the context's symmetric nature.
 
@@
 
-We recruited 360 US participants via Amazon Mechanical Turk in the speaker conditions, 180 in each condition.
-The participants were presented with a context similar to Figure \ref{exp_context}, together with a word denoting either of the shared features (color or shape).
-The participants were asked to imagine someone else was talking to them and he used the word to refer to one of the objects.
-Then they were asked to choose the object they thought that person was talking about.
+There were 600 participants in the listener conditions, 240 in the salience condition and 180 in each of the remaining two conditions.
+The numbers of participants choosing each object in each condition are shown in Table \ref{table:listener}.
 
@@
   \centering
-\begin{tabular}{c|ccc}
-   & \quad \textcolor{green!65}{\Large{$\blacksquare$}}&  \textcolor{green!65}{\Huge{$\bullet$}}& \textcolor{blue!65}{\Huge{$\bullet$}} \\ 
+\begin{tabular}{c|cccc}
+   & \quad \textcolor{green!65}{\Large{$\blacksquare$}}&  \textcolor{green!65}{\Huge{$\bullet$}}& \textcolor{blue!65}{\Huge{$\bullet$}} & Total\\ 
      \hline
- ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & $0$      \\
- ``Circle''   &  \quad $1$        &    $62$        & $117$
+ Salience   \quad  & \quad  $71$      \quad    &   $30$    \quad     & $139$ \quad     & $240$      \\
+ ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & $0$ \quad     & $180$     \\
+ ``Circle''   &  \quad $1$        &    $62$        & $117$ \quad     & $180$
   \end{tabular}
@@
 
-The numbers of participants choosing each object upon hearing each word are shown in Table \ref{table:listener}.
-Note that in either case, the object with both features shared (the green circle) is what the Gricean pragmatic account predicts to be the target.
-
-The result in Table \ref{table:listener} shows that when the message is ``Green'', listeners prefer the green circle which is predicted by the Gricean pragmatics, while they tend to stick to the blue circle which is more perceptually salient when the message is ``Circle''\footnote{Admittedly it is strange that one participant chose the square when the message was ``Circle''.
-However, since he passed the attention check we just include his data as it is.}.
+The result of the salience condition will be used as the empirical estimation of contextual salience.
+For the other two conditions, the object with both features shared (the green circle in the above table, but note that we counterbalanced colors) is what the Gricean pragmatic account predicts to be the target.
+The result in Table \ref{table:listener} shows that when the message is ``Green'', listeners prefer the green circle which is predicted by the Gricean pragmatics, while they tend to stick to the blue circle which is more perceptually salient when the message is ``Circle''.
 The behavioral patterns in both conditions are significantly different from uniform random choice, and they significantly differ from each other in respect of whether they conform to the predictions by Gricean pragmatic theory.

### 2013-10-09 01_intro.tex
intro/conclusion

@@
 We first introduce the simple referential communication game investigated in \cite{Frank} and the original RSA model proposed there.
-We then analyze a few design choices in this model and introduce their alternatives in game-theoretic pragmatics.
-Next we report the result of an experiment similar to the original one, and compare the predictions of different models to the new data.
-Finally we report on an additional experiment that suggests the sparsity of higher-order inference in actual human pragmatic reasoning.
+Next we analyze a few design choices in this model and introduce their alternatives in game-theoretic pragmatics.
+Finally we report on the result of an experiment similar to the original one, and compare the predictions of different models to the new data.

### 2013-10-09 06_conclusion.tex
intro/conclusion

@@
-Our main contributions are as follows: (1) We emphasized the distinction between belief, goal and action, systematically examined the RSA model, connected it to the game-theoretic approaches.
-(2) We tested the RSA model with a family of alternatives on the forced choice experiment and drew attention to the speaker's preference.
-Our data seems to suggest a different story about common knowledge --- what is commonly known is the Gricean pragmatic components, while both sides have his own private bias unknown to the other.
-(3) We explored the possibility of speakers doing second-order pragmatic inference and showed that it seems hard for them to do so.
+Our experiment and model comparison suggest the following: (1) on the conceptual level, 
+it is helpful to clarify the distinction between belief, goal and action, and by systematically examining each component in the formal model we can more explicitly spell out the underlying assumptions and formulate alternatives to be tested.
+(2) on the empirical level, we tested the RSA model with a family of alternatives on a forced-choice experiment and investigated the speaker's preference as well as the listener's perceptual salience.
+Our data seem to suggest a different story about common knowledge --- what is commonly known is the Gricean pragmatic components, while both sides have his own private bias unknown to the other.
 
-Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to further improvement in predictive power.
-Of course, more data from a wide range of further experiments are needed for a full understanding of pragmatic inference about referential expressions.
+Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to a better understanding of the phenomenon and further improvement in the formal model's predictive power.
+Of course, more empirical data are needed for a robust pragmatic theory of people's reasoning about referential expressions.

### 2013-10-14 06_conclusion.tex
cross-ref & conclusion

@@
-Our experiment and model comparison suggest the following: (1) on the conceptual level, 
-it is helpful to clarify the distinction between belief, goal and action, and by systematically examining each component in the formal model we can more explicitly spell out the underlying assumptions and formulate alternatives to be tested.
-(2) on the empirical level, we tested the RSA model with a family of alternatives on a forced-choice experiment and investigated the speaker's preference as well as the listener's perceptual salience.
-Our data seem to suggest a different story about common knowledge --- what is commonly known is the Gricean pragmatic components, while both sides have his own private bias unknown to the other.
+Our experiment and model comparison suggest the following: (1) On the conceptual level, 
+it is helpful to clarify the distinction between belief, goal and action, and the Bayesian framework provides us with a natural way to structure these components into a formal model.
+By examining each component systematically, we can explicitly spell out the underlying assumptions and formulate alternative hypotheses to be tested.
+(2) On the empirical level, we tested the RSA model with a family of variants motivated from the game-theoretic pragmatic perspectives on communication, and in particular, we investigated the speaker's preference as well as the listener's perceptual salience.
+Our data showed a more intricate picture about the roles various factors play in pragmatic reasoning about referential expressions --- what is commonly known is the Gricean pragmatic components, while either side have his own private bias unknown to the other.
+Also, an action-based view might better reflect the goal of communication, at least in the forced-choice experimental setting.
 

### 2013-10-23 01_intro.tex
Latest version

@@
 
-Despite the RSA model's theoretical advantage in that it provides us with quantitative predictions and its empirical success, if we analyze the design of the model, we will find several choices that are not \emph{prima facie} obvious and thus need further clarification.
+Despite the RSA model's theoretical advantage in that it provides us with quantitative predictions and its empirical success, if we analyze the design of the model, we will find several choices that are not \emph{prima facie} obvious and thus need further consideration and justification.
 These choices have their alternatives in the closely related game-theoretic approaches to pragmatics (e.g.
 \cite{Benz2007,Jager2013}), which have similar but slightly different conceptual motivations.
-Hence it is important to systematically compare the original RSA model with all these alternatives and test their predictions on empirical data, to gain a better understanding of the nature of human pragmatic inference.
+Hence it is important to systematically compare the original RSA model with all these alternatives and test their predictions on empirical data, to gain a better understanding of the relation between these models.
+Doing so will help illuminate the theoretical commitment and empirical implication of each design choice, which will enhance our understanding of the nature of human pragmatic inference.
 
@@
 We first introduce the simple referential communication game investigated in \cite{Frank} and the original RSA model proposed there.
-Next we analyze a few design choices in this model and introduce their alternatives in game-theoretic pragmatics.
-Finally we report on the result of an experiment similar to the original one, and compare the predictions of different models to the new data.
+Next we analyze a few design choices in this model and introduce their alternatives in game-theoretic pragmatics, emphasizing the distinction between \emph{belief}, \emph{goal} and \emph{action} and unifying various models as different interpretations of these three components.
+Then we report on the result of an experiment similar to the original one \cite{Frank}, and compare the predictions of different models to the new data.
+Finally, we discuss the implications of the results of the model comparison, pointing out that a more complex picture of pragmatic reasoning about referential expressions awaits further exploration.

### 2013-10-23 04_experiment.tex
Latest version

@@
 
-The positions and colors of the objects, as well as the orders of the candidate words in the speaker conditions, were counterbalanced.
+In order to minimize the effect of the confounding factors, we counterbalanced the positions and colors of the objects, as well as the orders of the candidate words in the speaker conditions.
+Thus the unique color in a context could be either blue (as in Fig.
+\ref{exp_context}) or green and the order of the objects could be any of the permutations\footnote{Due to the scale of our preliminary experiment we did not counterbalance the shape of the objects, i.e.
+a context always has one square and two circles.
+}.
 
@@
 The numbers of participants choosing each word in each condition are shown in Table \ref{table:speaker}.
-Note that when the target was the object with the unique shape (as in the first row of the table), the feature of shape (``Square'' in the example) should be the optimal utterance because the listener could uniquely identify it.
-Similarly when the target was the object with the unique color (the second row), the optimal utterance would be the feature of color (``Blue'').
+Note that when the target was the object with the unique shape (as in the first row of the table), the feature of shape (``Square'') should be the optimal utterance because the listener could uniquely identify it.
+Similarly when the target was the object with the unique color (the second row), the optimal utterance would be the feature of color (``Blue'' in this case).
 When the target was the object with both features shared, both features should be equally ambiguous because of the context's symmetric nature.

### 2013-10-23 06_conclusion.tex
Latest version

@@
 (2) On the empirical level, we tested the RSA model with a family of variants motivated from the game-theoretic pragmatic perspectives on communication, and in particular, we investigated the speaker's preference as well as the listener's perceptual salience.
-Our data showed a more intricate picture about the roles various factors play in pragmatic reasoning about referential expressions --- what is commonly known is the Gricean pragmatic components, while either side have his own private bias unknown to the other.
+Our data showed a more intricate picture about the roles various factors play in pragmatic reasoning about referential expressions.
+There is some correlation between the speaker's and the listener's rationality,  while either side has his own bias and is unaware of the other's.
 Also, an action-based view might better reflect the goal of communication, at least in the forced-choice experimental setting.
+Understanding the relation between the forced-choice and the betting paradigms would be an important next step.
 
 Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to a better understanding of the phenomenon and further improvement in the formal model's predictive power.
-Of course, more empirical data are needed for a robust pragmatic theory of people's reasoning about referential expressions.
+Of course, our preliminary experiment should only serve as a start and more empirical data are needed for a robust pragmatic theory of people's reasoning about referential expressions.

### 2013-10-23 main.tex
micha push test
model comparison speaker
ModComp nicer tables and added references
plots updated, packages added to main.tex
Latest version

@@
 \usepackage{graphicx}
+
+\usepackage[table]{xcolor}
 
@@
 %\usepackage{lingmacros}
+
+%%% Micha's stuff
+\usepackage{booktabs}
+\usepackage{todonotes}
+
 % ======== Definitions ========
@@
 Despite its theoretical advantage and empirical success, there are still several conceptual issues in the Bayesian Rational Speech Act (RSA) model that need further clarification.
-In this paper, we analyze a few design choices made in RSA that have intuitively equally plausible alternatives motivated from related game-theoretic approaches to pragmatics, and compare the original RSA model with these alternatives to empirical data.
-The result suggests that while RSA is better than most other variations, there is still room for improvement.
+In this paper, we analyze a few design choices made in RSA, focusing on its \emph{belief}, \emph{goal} and \emph{action} components and introduce some intuitively equally plausible alternatives motivated from related game-theoretic approaches to pragmatics as different interpretations of these three components.
+Then we compare the original RSA model with these alternatives to new empirical data and the result suggests a more complex picture of pragmatic reasoning about referential expressions than the RSA depicts.
+While there is a correlation between the speaker's and the listener's pragmatic reasoning that may be seen as common rationality, either side has its own bias that the other does not take into account.
 \end{abstract}

### 2013-10-28 main.tex
revised abstract

--- 2013-09-18 main.tex
@@
 % ======= Packages ======
-
+\usepackage{etex}
 %    ### General ###
@@
 
+\usepackage[table]{xcolor}
+
 %    ### Game Theory ###
@@
    \usetikzlibrary{shapes.geometric}
-
+\usepackage{pgfplots}
 %    ### Linguistics ###
@@
 \begin{abstract}
-Recent developments in Bayesian experimental pragmatics have received much attention, as it provides us with a quantitative model formalizing core concepts in traditional pragmatic theories and the predictions it makes fit empirical data nicely.
-Despite its theoretical advantage and empirical success, there are still several conceptual issues in the Bayesian Rational Speech Act (RSA) model that need further clarification.
-In this paper, we analyze a few design choices made in RSA that have intuitively equally plausible alternatives motivated from related game-theoretic approaches to pragmatics, and compare the original RSA model with these alternatives to empirical data.
-The result suggests that while RSA is better than most other variations, there is still room for improvement.
+Recent developments in Bayesian experimental pragmatics have received much attention.
+The Rational Speech Act (RSA) model formalizes core concepts in traditional pragmatic theories quantitatively and makes predictions that fit empirical data nicely.
+In this paper, we analyze the RSA model and its relation to the closely related game-theoretic approaches to pragmatics, by spelling out its \emph{belief}, \emph{goal} and \emph{action} components.
+We introduce some alternatives motivated from the game-theoretic models and compare them with the RSA model to new experimental data.
+The result suggests that the RSA model could be adapted and extended to improve its predictive power.
+
 \end{abstract}

### 2013-10-30 01_intro.tex
rephrased intro, amended layout

--- 2013-10-09 01_intro.tex
@@
-Human language communication is efficient, in that people need not always say every detail in order to be understood.
-Rather, speakers only say what is most relevant, and listeners can seemingly effortlessly grasp the intended meaning beyond what is literally said.
-This ability to do pragmatic inference has been long studied and the \emph{conversational implicature} theory by Grice \cite{Grice} is one of the most prominent theories in the field.
-However, since it is hard to formalize the \emph{Cooperative Principle} and the \emph{Conversational Maxims}, Grice's theory is not precise enough to give empirically testable predictions, especially quantitative predictions.
+Human language communication is efficient, in that people need not
+always say every detail in order to be understood. Rather, speakers
+often only say what is most relevant, and listeners can often
+seemingly effortlessly grasp the intended meaning beyond what is
+literally said. This ability to do pragmatic inference has been long
+studied and the \emph{conversational implicature} theory by Grice
+\cite{Grice} is one of the most prominent theories in the
+field. However, since it is hard to formalize the \emph{Cooperative
+  Principle} and the \emph{Conversational Maxims}, Grice's theory is
+not precise enough to give empirically testable predictions,
+especially quantitative predictions.
 
-The Bayesian \emph{Rational Speech Act} (RSA) model attempts to address this issue by using information-theoretic notions together with the Bayesian inference framework.
-It has been shown that the model could yield quantitative predictions that are highly correlated to actual human judgments \cite{Frank}.
+The Bayesian \emph{Rational Speech Act} (RSA) model attempts to
+address this issue by using information-theoretic notions together
+with the Bayesian inference framework
+\cite{Frank2009,Frank,Bergen2012,GoodmanStuhlmuller2013:Knowledge-and-I}. It
+has been shown that the model could yield quantitative predictions
+that are highly correlated to actual human judgments
+\cite{Frank,GoodmanStuhlmuller2013:Knowledge-and-I}.
 
-Despite the RSA model's theoretical advantage in that it provides us with quantitative predictions and its empirical success, if we analyze the design of the model, we will find several choices that are not \emph{prima facie} obvious and thus need further clarification.
-These choices have their alternatives in the closely related game-theoretic approaches to pragmatics (e.g.
-\cite{Benz2007,Jager2013}), which have similar but slightly different conceptual motivations.
-Hence it is important to systematically compare the original RSA model with all these alternatives and test their predictions on empirical data, to gain a better understanding of the nature of human pragmatic inference.
+Despite the RSA model's theoretical promise and empirical success, if
+we analyze the design of the model, we will find several choices that
+are not \emph{prima facie} obvious and might thus need further
+consideration and justification. These choices have their alternatives
+in closely related game-theoretic approaches to pragmatics
+(e.g. \cite{Benz2007,Jager2013}), which have similar but slightly
+different conceptual motivations. Hence it is important to
+systematically compare the original RSA model with all these
+alternatives and test their predictions on empirical data, so as to
+gain a better understanding of the relation between these
+models. Doing so will help illuminate the theoretical commitments and
+empirical implications of each design choice, which will enhance our
+understanding of the nature of human pragmatic inference.
 
-The outline of the paper is as follows.
-We first introduce the simple referential communication game investigated in \cite{Frank} and the original RSA model proposed there.
-Next we analyze a few design choices in this model and introduce their alternatives in game-theoretic pragmatics.
-Finally we report on the result of an experiment similar to the original one, and compare the predictions of different models to the new data.
+The outline of the paper is as
+follows. Section~\ref{sec:refer-games-rsa} introduce the referential
+communication games investigated by \cite{Frank} and the original RSA
+model proposed there. Section~\ref{sec:alternatives} analyzes a few
+design choices in this model and introduces their alternatives in game
+theoretic pragmatics, emphasizing the distinction between
+\emph{belief}, \emph{goal} and \emph{action} and unifying various
+models as different interpretations of these three
+components. Section~\ref{sec:experiment} reports on the results of an
+experiment similar to the one of \cite{Frank}. In
+Section~\ref{sec:model-comparison}, we compare the predictions of
+different models of our new data. Finally, we discuss the implications
+of the results of the model comparison in
+Section~\ref{sec:conclusion}, concluding that the original RSA model
+can be improved on, in particular by taking speaker preferences into
+account, but we also point out that a more complex picture of
+probabilistic pragmatic reasoning about referential expressions awaits
+further exploration.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "main"
+%%% TeX-PDF-mode: t
+%%% End:

### 2013-10-30 main.tex
rephrased intro, amended layout

@@
 
-\title{Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning}
-\author{Michael Franke \and Ciyang Qing}
-\institute{Institute for Logic, Language and Computation}
+\title{Variations on a Bayesian Theme: Comparing Bayesian Models of
+  Referential Reasoning}
+\titlerunning{Variation on a Bayesian Theme}
+\author{Ciyang Qing \and Michael Franke\thanks{Thanks to \dots % Judith Degen,
+  % Noah Goodman, Daniel Lassiter [?], audience of the workshop, Henk
+  % & Hans-Christian for organizing, reviewers \dots 
+}}
+\institute{Institute for Logic, Language and Computation\\Universtity
+  of Amsterdam, The Netherlands}
 
@@
 \begin{abstract}
-Recent developments in Bayesian experimental pragmatics have received much attention.
-The Rational Speech Act (RSA) model formalizes core concepts in traditional pragmatic theories quantitatively and makes predictions that fit empirical data nicely.
-In this paper, we analyze the RSA model and its relation to the closely related game-theoretic approaches to pragmatics, by spelling out its \emph{belief}, \emph{goal} and \emph{action} components.
-We introduce some alternatives motivated from the game-theoretic models and compare them with the RSA model to new experimental data.
-The result suggests that the RSA model could be adapted and extended to improve its predictive power.
-
+  Recent developments in Bayesian experimental pragmatics have
+  received much attention. The Rational Speech Act (RSA) model
+  formalizes core concepts of traditional pragmatic theories
+  quantitatively and makes predictions that fit empirical data
+  nicely. In this paper, we analyze the RSA model and its relation to
+  closely related game theoretic approaches, by spelling out its
+  \emph{belief}, \emph{goal} and \emph{action} components. We
+  introduce some alternatives motivated from the game theoretic
+  tradition and compare models incorporating these alternatives
+  systematically to the original RSA model, using Bayesian model
+  comparison, in terms of their ability to predict relevant empirical
+  data. The result suggests that the RSA model could be adapted and
+  extended to improve its predictive power, in particular by taking
+  speaker preferences into account. 
 \end{abstract}
@@
 \section{Referential Games and the RSA model}
+\label{sec:refer-games-rsa}
 
@@
 \section{Alternatives in Design Choices}
+\label{sec:alternatives}
 
@@
 \section{The Experiment}
+\label{sec:experiment}
 
@@
 \section{Model Comparison}
+\label{sec:model-comparison}
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2013-10-30 01_intro.tex
Sections 2,3 proof

@@
 in closely related game-theoretic approaches to pragmatics
-(e.g. \cite{Benz2007,Jager2013}), which have similar but slightly
+\cite{Rabin1990:Communication-b,Stalnaker:SayingMeaningCredibility,BenzvanRooijOptimalAssertions2007,Franke2011:Quantity-Implic,Jager2013:Rationalizable-}, which have similar but slightly
 different conceptual motivations. Hence it is important to

### 2013-10-30 02_RSA.tex
rephrased intro, amended layout
Sections 2,3 proof

@@
-Figure \ref{context} is an example of the referential game used in studying human pragmatic inference \cite{Frank}.
-In a referential game there is a set of objects called the \emph{context}.
-The speaker in the game is asked to refer to one of the objects (the \emph{target}) by uttering an expression (typically a word) to the listener.
-The listener, who has no idea which object is the target, needs to recover the speaker's intended referent based on the speaker's utterance.
+Referential games are confined interactive reasoning tasks used to
+study pragmatic inference in a controlled environment \cite{Frank}. A
+referential game consists of a set of objects called a
+\textit{context}. A context contains different shapes of different
+colors in various arrangements, such as shown in Figure~\ref{context}.
+The speaker in the game is asked to refer to one of the objects (the
+\emph{target}) by uttering an expression (typically a single word
+denoting a feature, i.e., color or shape, of the target) to the
+listener. The listener, who has does not know which object is the
+target, needs to recover the target based on the speaker's choice of
+expression.
 
 \begin{figure}[h] 
-  \centering
-  \subfigure[The context]{\label{context}
+  \centering \subfigure[Example context: green square and circle, and
+  blue circle.]{\label{context}
   \begin{tikzpicture}
@@
   \end{tikzpicture}
-  }  
-  \hspace{5mm}
-  \subfigure[The vocabulary and the truth table]{\label{VaTT}
+}\hspace{5mm} \subfigure[Vocabulary and truth table]{\label{VaTT}
   \begin{tikzpicture}
@@
 
-Let us first try to play this game ourselves, to get a better sense of what it is about and why it is relevant to human pragmatic reasoning.
-For simplicity let us assume that the utterances that the speaker is allowed to use are from a \emph{vocabulary} which is commonly known between the speaker and the listener.
-The content of the vocabulary and the (truth-conditional/literal) meaning of each word are shown in Figure \ref{VaTT}.
-Suppose we play as the speaker and the target is $t_1$, the green square.
-We can either use ``Square'' or ``Green'' to literally truthfully refer to it, but it seems more prudent to use ``Square'', as there is only one square in the context and thus the listener can easily identify it, whereas there are two green objects, which makes ``Green'' ambiguous.
-In terms of the Gricean \emph{Maxim of Quantity}, we should use ``Square'' because it is as maximally informative as is required.
-Similarly, we should use ``Blue'' to refer to $t_3$, the blue circle.
-However,  using ``Blue'' to refer to the blue circle might intuitively sound a little unnatural, as color terms are usually used as adjectives\footnote{They are used as nouns mostly to refer to the colors themselves.}, while we usually use nouns to refer to concrete objects.
-This inclination becomes more evident when we want to refer to $t_2$, the green circle.
-While ``Green'' and ``Circle'' are equally ambiguous in this case, we might nevertheless prefer the latter.
+Let us first try to play this game ourselves, to get a better sense of
+what it is about and why it is relevant to human pragmatic
+reasoning. For simplicity let us assume that the utterances that the
+speaker is allowed to use are from a \emph{vocabulary} which is
+commonly known between the speaker and the listener. The content of
+the vocabulary and the (truth-conditional/literal) meaning of each
+word are shown in Figure \ref{VaTT}. Suppose we play as the speaker
+and the target is $t_1$, the green square. We can either use
+``Square'' or ``Green'' to truthfully refer to it, but it seems more
+prudent to use ``Square'', as there is only one square in the context
+and thus the listener can easily identify it, whereas there are two
+green objects, which makes ``Green'' ambiguous. In terms of the
+Gricean \emph{Maxim of Quantity}, we should use ``Square'' because it
+is more informative than ``Green'' in the given context; the surplus
+informativity moreover seems relevant. Similarly, we should use
+``Blue'' to refer to $t_3$, the blue circle. However, using ``Blue''
+to refer to the blue circle might intuitively sound a little
+unnatural, as color terms are usually used as adjectives\footnote{They
+  are used as nouns mostly to refer to the colors themselves.}, while
+we usually use nouns to refer to concrete objects. This inclination
+becomes more evident when we want to refer to $t_2$, the green
+circle. While ``Green'' and ``Circle'' are equally ambiguous in this
+case, we might nevertheless prefer the latter.
 
-Now let us turn to play as the listener.
-If we hear ``Square'' or ``Blue'', we easily know the intended referent as there is no ambiguity, but what if we hear ``Circle'' (or ``Green'')?
-There are two circles in the context that we need to choose from.
-On the one hand, the blue circle, having the unique color in the context, seems to be perceptually dominant and thus easily captures our attention.
-On the other hand, from the previous analysis we know that if the blue circle were the intended referent, the speaker could have chosen ``Blue'' which is not ambiguous and thus more informative.
-Hence the listener needs to balance two sources of information, i.e.
-the (presumably subconscious) perceptual salience of different objects and the rational expectation of the likelihood of the speaker making the utterance for each object.
-The latter line of reasoning is crucial to the classic Gricean account of scalar implicatures, and the major challenges in terms of formal modeling are how to quantify  notions such as informativeness and how different pieces of information should be integrated.
+Now let us turn to play as the listener. If we hear ``Square'' or
+``Blue'', we easily know the intended referent as there is no
+ambiguity, but what if we hear ``Circle'' (or ``Green'')? There are
+two circles in the context that we need to choose from. On the one
+hand, the blue circle, having the unique color in the context, seems
+to be perceptually dominant and thus easily captures our attention. On
+the other hand, from the previous analysis we know that if the blue
+circle were the intended referent, the speaker could have chosen
+``Blue'' which is not ambiguous and thus more informative. Hence the
+listener needs to balance two sources of information, i.e., the
+(presumably subconscious) perceptual salience of different objects and
+the rational expectation of the likelihood of the speaker making the
+utterance for each object. The latter line of reasoning is crucial to
+the classic Gricean account of scalar implicatures, and the major
+challenges in terms of formal modeling are how to quantify notions
+such as informativeness and how different pieces of information should
+be integrated.
 
-The RSA model addresses the problems by using information-theoretic concepts to measure informativeness and Bayesian inference to integrate different sources of information \cite{Frank}.
-
-In order to measure the informativeness of an utterance, the RSA model starts with the literal listener who upon receiving an utterance $m$ does a conditioning on its literal meaning:
+The RSA model addresses the problems by using information-theoretic
+concepts to measure informativeness and Bayesian inference to
+integrate different sources of information
+\cite{Frank2009,Frank,Bergen2012,GoodmanStuhlmuller2013:Knowledge-and-I}. In
+order to measure the informativeness of an utterance, the RSA model
+starts with the literal listener who upon receiving an utterance $m$
+does a conditioning on its literal meaning:
 \begin{equation}\label{Bayesian_literal}
@@
 \end{equation}
-where $\mathcal{U}$ is a uniform distribution over all the possible referents.
-Now the informativeness of utterance $m$ for the intended referent $t$ can be measured as the negative Kullback-Leibler divergence of the induced literal listener's belief $\rho_0$ from the speaker's own belief $\delta_t$:
+where $\mathcal{U}$ is a uniform distribution over all the possible
+referents. The informativeness of utterance $m$ for the intended
+referent $t$ can be measured as the negative Kullback-Leibler
+divergence of the induced literal listener's belief $\rho_0$ from the
+speaker's own belief $\delta_t$:
 \begin{equation} \label{Bayesian_sender_informativeness}
@@
 \end{equation}
-where $\delta_t$ is a delta distribution with all the probability mass on $t$, as the speaker knows her intended referent:
+where $\delta_t$ is a delta distribution with all probability mass on
+target object $t$, as the speaker knows her intended referent:
 \begin{equation} \label{Bayesian_sender_delta}
@@
 \end{equation}
-where $\lambda_\mathrm{S}$ is a parameter measuring the speaker's degree of rationality, i.e.
-to what extent the speaker sticks to the strict optimum.
-The cost term is used to encode preference in different utterances, be it about the utterances' lengths or syntactic categories.
+where $\lambda_\mathrm{S}$ is a parameter measuring the speaker's
+degree of rationality, i.e., to what extent the speaker sticks to the
+strict optimum. The cost term is used to encode preference in
+different utterances, be it about the utterances' lengths or syntactic
+categories.
 
@@
 
-The pragmatic listener in the RSA model, upon receiving the utterance $m$, does a Bayes update on his prior belief $\mathcal{S}(t)$ by using his knowledge of the speaker's behavior (\ref{Bayesian_speaker}):
+The pragmatic listener in the RSA model, upon receiving the utterance
+$m$, performs a Bayesian update on his prior belief $\mathcal{S}(t)$
+by using an estimate of the speaker's behavior
+(\ref{Bayesian_speaker}):
 \begin{equation} \label{Bayesian_rec_update}
@@
 \end{equation}
-The Bayes rule helps naturally integrate the perceptual salience of each object, which is treated as the prior $\mathcal{S}(t)$ and can be empirically measured, with listener's expectation of the speaker being informative, which is incorporated as the likelihood, thus addressing the previously mentioned challenge of balancing different sources of information.
-Setting $\lambda=1$ and $\mbox{Cost}(m)=0$ for all $m$, \cite{Frank} obtained a highly significant correlation between the model prediction and the experiment data.
+Bayes' rule naturally integrates the perceptual salience of each
+object, which is treated as the prior $\mathcal{S}(t)$ and can be
+empirically measured, with listener's expectation of the speaker being
+informative, which is incorporated as the likelihood, thus addressing
+the previously mentioned challenge of balancing different sources of
+information. Setting $\lambda=1$ and $\mbox{Cost}(m)=0$ for all $m$,
+\cite{Frank} obtained a highly significant correlation between the
+model prediction and their experiment data on actual speaker and
+listener behavior gathered from referential games of varying
+complexity.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "main"
+%%% TeX-PDF-mode: t
+%%% End:

### 2013-10-30 03_alternatives.tex
cross-ref & conclusion
rephrased intro, amended layout
Sections 2,3 proof

@@
-The RSA model has the theoretical advantage in that it is quantitative, and it has been shown to be able to fit the empirical data well.
-However, if we want to gain more insight about communication involving referential expressions from its success, it will be worthwhile to examine the model carefully to pin down the major components, spell out the design choices and the underlying assumptions, test their contribution to the predictive power of the model through comparison with the alternatives and scrutinize their implications thereof.
-In this section, we will carry out a conceptual analysis of the RSA model by comparing it to the closely related game-theoretic approaches to pragmatics, to define a class of alternative models.
-We will then compare the predictions of these models to the empirical data in later sections.
+The RSA model has a theoretical advantage over traditional formal
+pragmatic theories in that it provides quantitative predictions, and
+it has been shown to fit relevant empirical data well. However, if we
+want to learn more about pragmatic reasoning about referential
+expressions, it will be worthwhile to examine RSA carefully to pin
+down its major components, to spell out its main design choices and
+their underlying assumptions, and to test their contribution to the
+predictive power of the model through statistical comparison with
+natural alternatives. In this section, we will therefore compare the
+RSA model to closely related game theoretic approaches, that likewise
+assume that speakers and listeners form cascading beliefs about mutual behavior
+and seek to optimize their behavior based on these beliefs
+\cite{Rabin1990:Communication-b,Stalnaker:SayingMeaningCredibility,BenzvanRooijOptimalAssertions2007,Franke2011:Quantity-Implic,Jager2013:Rationalizable-}.\footnote{See
+\cite{FrankeJager2013:Pragmatic-Back-} for overview and comparison
+of game theoretic and Bayesian models.} Doing so
+let's us describe a class of potential alternatives to RSA. We will
+then compare the predictive power of these given relevant empirical
+data in later sections.
 
-First and foremost, there are three closely related notions, i.e.
-\emph{belief}, \emph{goal} and \emph{action}, that play a crucial role in the RSA as well as game-theoretic models.
-The relation and distinction between them can be best illustrated by looking at (\ref{Bayesian_speaker}):
+There are three closely related notions, i.e., \emph{belief},
+\emph{goal} and \emph{action}, that play a crucial role in RSA as well
+as game theoretic models. The relation and distinction between them
+can be best illustrated by looking at (\ref{Bayesian_speaker}),
+repeated here:
 $$\sigma(m \mid t) \propto \exp(\lambda_\mathrm{S} \cdot (\log \mathcal{U}(t\mid \intp{m})-\mbox{Cost}(m))) \enspace .$$
-The term $\mathcal{U}(t\mid \intp{m})$ is the speaker's \emph{belief} about the listener's interpretation of the utterance.
-The expected utility $\mbox{U}(m,t)=\log \mathcal{U}(t\mid \intp{m})-\mbox{Cost}(m)$ describes the speaker's \emph{goal} of communication, i.e.
-inducing a belief that is as close as his own with minimal effort.
-The production rule $\sigma(m \mid t)$ on the left-hand side specifies the speaker's \emph{action}, as it defines the probability the speaker actually chooses an action, i.e.
-uttering a particular expression.
-The soft-max rule connects these notions, as a (sub-)rational agent's action depends on his belief and goal.
-While the RSA and game-theoretic models all share this general architecture in their designs, they vary in the specific interpretations of the three ingredients, which reflect different views and emphasis on language and communication and give rise to a series of interesting and important empirical questions.
-Again, let us continue with the speaker model (\ref{Bayesian_speaker}) to illustrate some points of divergence and formally define the corresponding alternatives.
+The term $\mathcal{U}(t\mid \intp{m})$ is the speaker's \emph{belief}
+about the listener's interpretation of the utterance. The expected
+utility $\mbox{U}(m,t)=\log \mathcal{U}(t\mid
+\intp{m})-\mbox{Cost}(m)$ describes the speaker's \emph{goal} of
+communication, i.e., inducing a belief that is as close as possible to
+her own with minimal effort. The production rule $\sigma(m \mid t)$ on
+the left-hand side specifies the speaker's \emph{action}, as it
+defines the probability with which, according to RSA, speakers would
+choose a particular expression. The soft-max rule connects these
+notions, as a (sub-)rational agent's action depends on his belief and
+goal. While the RSA and game theoretic models all share this general
+architecture in their designs, they vary in the specific
+interpretations of the three ingredients, which reflect different
+views and emphasis on language and communication and give rise to a
+series of interesting and important empirical questions. Again, let us
+continue with the speaker model (\ref{Bayesian_speaker}) to illustrate
+some points of divergence and formally define the corresponding
+alternatives.
 
-First of all, although \cite{Frank} claims that the empirically measured perceptual salience is used to capture the notion of common knowledge between the speaker and the listener, it turns out that the speaker's production rule (\ref{Bayesian_speaker}) simply does not take the listener's salience prior $\mathcal{S}(t)$ into account in her literal listener model (\ref{Bayesian_literal}).
-This makes it unclear in what sense the speaker knows the listener's perceptual salience.
-A natural variant, which is used in some game-theoretic models, would be to replace the literal listener's uniform distribution $\mathcal{U}(t)$ in (\ref{Bayesian_literal}) with the salience prior $\mathcal{S}(t)$.
-This will lead to the alternative production rule:
+First of all, although it is prima facie reasonable to hypothesize, as
+\cite{Frank} do, that the empirically measured perceptual salience
+$\mathcal{S}(\cdot)$ is common knowledge between speaker and listener,
+it does not actually affect RSA's production rule
+(\ref{Bayesian_speaker}) at all. This makes it unclear in what sense
+the speaker can be said to know the listener's perceptual salience,
+since neither his beliefs nor actions depend on it. A natural variant,
+which is used in some game theoretic models, would replace the literal
+listener's uniform distribution $\mathcal{U}(t)$ in
+(\ref{Bayesian_literal}) with the salience prior
+$\mathcal{S}(t)$. This leads to the alternative production rule:
 \begin{equation} \label{Bayesian_speaker_salience}
@@
 
-Secondly, the informativeness of an utterance $m$, which is a crucial part of the communication goal, is measured in the RSA model in terms of how close the induced belief of the literal listener is from that of the speaker's own belief.
-This means that the RSA model sees the goal of communication as conveying belief.
-While it is normally true that language does convey the belief of the speaker, it is questionable at least in this referential scenario whether letting the listener form a proper belief is the ultimate goal of the communication.
-After all, if the speaker wants to refer to something, it seems that in the end what matters is whether the listener actually picks out the intended referent successfully (e.g., when the speaker wants the listener to pass something).
-We might call this view \emph{action-oriented}, in contrast to the \emph{belief-oriented} view of communication which the RSA model adopts, as it interprets the goal of communication as invoking the intended action rather than forming an accurate belief.
-\footnote{Note that the two views are definitely intimately related, despite our emphasis on their differences here.
-An accurate belief is usually helpful to choosing a desired action.
-In fact, forming a belief can itself be seen as an action that the listener performs, which might well be all that the speaker cares about in many everyday situations, e.g.
-casual talks or responses to inquiries of information.
-Also, the relevant actions might vary greatly across situations which makes belief formation a practical alternative to serve as the goal.
-Nevertheless, they are still essentially different views in spite of the fact that they often coincide in practice.
-The referential communication scenario we investigate in this paper reflects the distinction.} Thus, according to this view, the informativeness of a message would be measured as the probability of the listener choosing the intended referent.
-Formally speaking, we have the action-oriented speaker model
+Secondly, RSA measures the informativeness of an utterance $m$, which
+is a crucial part of the communicative goal, in terms of how close the
+induced belief of the literal listener is from the speaker's own
+belief. This means that the RSA model sees the goal of communication
+as conveying belief. While it is normally true that language does
+convey the belief of the speaker, it is questionable at least in this
+referential scenario whether letting the listener form a proper belief
+is the ultimate goal of the communication. After all, if the speaker
+wants to refer to something, it seems that in the end what matters is
+whether the listener actually picks out the intended referent
+successfully (e.g., when the speaker wants the listener to pass
+something). This view is inherent in game theoretic approaches where
+agents' beliefs are backed up by explicit formulations of their
+utilities. We might call this latter view \emph{action-oriented}, in
+contrast to the \emph{belief-oriented} view of communication which the
+RSA model adopts, as it interprets the goal of communication as
+invoking the intended action rather than forming an accurate
+belief.\footnote{Note that the two views are clearly intimately
+  related, despite our emphasis on their differences here. An accurate
+  belief is usually helpful for choosing a desired action. In fact,
+  forming a belief can itself be seen as an action that the listener
+  performs, which might well be all that the speaker cares about in
+  many everyday situations, e.g., casual talks or responses to
+  inquiries of information. Also, the relevant actions might vary
+  greatly across situations which makes belief formation a practical
+  alternative to serve as the goal. Nevertheless, they are still
+  essentially different views in spite of the fact that they often
+  coincide in practice. The referential communication scenario we
+  investigate in this paper reflects the distinction.} Thus, according
+to this view, the informativeness of a message would be measured as
+the probability of the listener choosing the intended
+referent. Formally speaking, we have the action-oriented speaker model
 
@@
 \end{equation}
-where $\mathcal{U}$ is the uniform prior and $\mathcal{S}$ is the salience prior.
-For example, in the original RSA model, the speaker does not take listener's salience prior into account and he has a belief-oriented goal of communication, thus it will be denoted as $\sigma_{\mathrm{b}\mathcal{U}}$
+where $\mathcal{U}$ is the uniform prior and $\mathcal{S}$ is the
+salience prior. For example, in the original RSA model, the speaker
+does not take listener's salience prior into account and he has a
+belief-oriented goal of communication. Thus it will be denoted as
+$\sigma_{\mathrm{b}\mathcal{U}}$.
 
-Finally, the original RSA speaker model in \cite{Frank} has $\mbox{Cost}(m)=0$ for all $m$, which means that the potential speaker preference in different utterances is not taken into account.
-As we previously pointed out, our intuition seems to suggest a preference for nouns over adjectives as referential expressions.
-Since it is an empirical question whether, to what direction or to what extent such a preference exists, we leave open all the possibilities.
-Technically speaking, since only the difference in costs matters, we define the cost function using a constant $c\in \mathbb{R}$
+Finally, the original RSA speaker model of \cite{Frank} has
+$\mbox{Cost}(m)=0$ for all $m$, which means that the potential speaker
+preference in different utterances is not taken into account (but not
+so in, e.g., \cite{Bergen2012}). As we previously pointed out, our
+intuition seems to suggest a preference for nouns over adjectives as
+referential expressions.\footnote{The rich tradition on referential
+  expression generation \cite{KramerDeemter2012:Computational-G} has
+  also identified the need to take speaker preferences into account,
+  and \cite{GattGompel2013:Are-we-Bayesian} criticize RSA on these
+  grounds as well.} Since it is an empirical question whether, to what
+direction or to what extent such a preference exists, we leave open
+all the possibilities. Technically speaking, since only the difference
+in costs matters, we define the cost function using a constant $c\in
+\mathbb{R}$
 \begin{equation} \label{cost}
@@
 
-Let us start from the relatively clear part.
-As pointed out previously, the likelihood term $\sigma(m \mid t)$ is the listener's belief about how the speaker behaves.
-One natural question is to what extent the listener's belief correlate with the actual production.
-We thus treat the speaker's production term $\sigma(m \mid t)$ in the listener's model as a parameter, making the association with a belief about production of the listener model explicit:
+Let us start from the relatively clear part. As pointed out
+previously, the likelihood term $\sigma(m \mid t)$ is the listener's
+belief about how the speaker behaves. One natural question is to what
+extent the listener's belief correlates with the actual production. We
+thus treat the speaker's production term $\sigma(m \mid t)$ in the
+listener's model as a parameter, making the association with a belief
+about production of the listener model explicit:
 \begin{equation} \label{listener-production}
@@
 \end{equation}
-Note that the speaker's production rule has two parameters $\lambda_\mathrm{S},c$ which are also be included in the above specification of the listener model.
+Note that the speaker's production rule has two parameters
+$\lambda_\mathrm{S},c$ which are also included in the above
+specification of the listener model.
 
@@
 
-Finally, since the RSA model adopts a belief-oriented view on the goal of communication, its listener model only consists of the belief about the intended referent after hearing the utterance, obtained by a Bayes update.
-However, according to the action-oriented view, this is not quite the end of the story.
-The listener often needs to decide the exact referent of a referential expression.
-(Again, consider the case in which the listener is asked to pass something.) Hence in such cases the listener will choose an object that he (soft-)maximally believes to be the intended referent.
-Formally speaking, the action-oriented listener model would be
+Finally, since the RSA model adopts a belief-oriented view on the goal
+of communication, its listener model only consists of the belief about
+the intended referent after hearing the utterance, obtained by a Bayes
+update. However, according to the action-oriented view, this is not
+quite the end of the story. The listener often needs to decide the
+exact referent of a referential expression. (Again, consider the case
+in which the listener is asked to pass something.) Hence in such cases
+the listener will choose an object by, essentially, (soft-)maximizing
+over his beliefs. Formally, the action-oriented listener model
+becomes:
 \begin{equation} \label{listener-action}
@@
 \end{equation}
-where $v\in\{\mathcal{U},\mathcal{S}\}$, $\lambda_\mathrm{L}$ is the parameter measuring the listener's degree of rationality, and $\rho_{\mathrm{b}v}$ is the belief-oriented model that does the Bayes update:
+where $v\in\{\mathcal{U},\mathcal{S}\}$, $\lambda_\mathrm{L}$ is the
+parameter measuring the listener's degree of rationality, and
+$\rho_{\mathrm{b}v}$ is the belief-oriented model that does the
+Bayesian update:
 \begin{equation} \label{listener-belief}
@@
 For instance, the original RSA listener model is a belief-oriented one with the perceptual salience as prior, whose belief about the speaker is $\sigma_{\mathrm{b}\mathcal{U}}$, hence it is denoted as $\rho_{\mathrm{b}\mathcal{S}}(\sigma_{\mathrm{b}\mathcal{U}})$.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "main"
+%%% TeX-PDF-mode: t
+%%% End:

### 2013-10-30 04_experiment.tex
rephrased intro, amended layout
Sections 2,3 proof

@@
 Thus the unique color in a context could be either blue (as in Fig.
-\ref{exp_context}) or green and the order of the objects could be any of the permutations\footnote{Due to the scale of our preliminary experiment we did not counterbalance the shape of the objects, i.e.
-a context always has one square and two circles.
+\ref{exp_context}) or green and the order of the objects could be any of the permutations\footnote{Due to the scale of our preliminary experiment we did not counterbalance the shape of the objects, i.e., a context always has one square and two circles.
 }.
@@
 There are four major designs in our own experiment that differ from the original experiment in \cite{Frank}.
-First of all, the original experiment used the betting paradigm, i.e.
-participants were asked to bet over a range of options and they were instructed that the amount of money that is bet on an option should reflect their confidence that the option is correct.
+First of all, the original experiment used the betting paradigm, i.e., participants were asked to bet over a range of options and they were instructed that the amount of money that is bet on an option should reflect their confidence that the option is correct.
 Even though the betting paradigm has the merit of providing us with graded reponses from each individual participant, the caveat is that it is unclear whether it measures the belief or the action.
@@
 The behavioral patterns in both conditions are significantly different from uniform random choice, and they significantly differ from each other in respect of whether they conform to the predictions by Gricean pragmatic theory.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "main"
+%%% TeX-PDF-mode: t
+%%% End:

### 2013-10-30 06_conclusion.tex
rephrased intro, amended layout
Sections 2,3 proof

@@
-Our experiment and model comparison suggest the following: (1) On the conceptual level, 
-it is helpful to clarify the distinction between belief, goal and action, and the Bayesian framework provides us with a natural way to structure these components into a formal model.
-By examining each component systematically, we can explicitly spell out the underlying assumptions and formulate alternative hypotheses to be tested.
-(2) On the empirical level, we tested the RSA model with a family of variants motivated from the game-theoretic pragmatic perspectives on communication, and in particular, we investigated the speaker's preference as well as the listener's perceptual salience.
-Our data showed a more intricate picture about the roles various factors play in pragmatic reasoning about referential expressions.
-There is some correlation between the speaker's and the listener's rationality,  while either side has his own bias and is unaware of the other's.
-Also, an action-based view might better reflect the goal of communication, at least in the forced-choice experimental setting.
-Understanding the relation between the forced-choice and the betting paradigms would be an important next step.
+Our experiment and model comparison suggest the following: (1) On the
+conceptual level, it is helpful to clarify the distinction between
+belief, goal and action, and the Bayesian framework provides us with a
+natural way to structure these components into a formal model. By
+examining each component systematically, we can explicitly spell out
+the underlying assumptions and formulate alternative hypotheses to be
+tested. (2) On the empirical level, we tested the RSA model with a
+family of variants motivated from the game theoretic pragmatic
+perspectives on communication, and in particular, we investigated the
+speaker's preference as well as the listener's perceptual
+salience. Our data showed a more intricate picture about the roles
+various factors play in pragmatic reasoning about referential
+expressions. There is some correlation between the speaker's and the
+listener's rationality, while either side has his own bias and is
+unaware of the other's. Also, an action-based view might better
+reflect the goal of communication, at least in the forced-choice
+experimental setting. Understanding the relation between the
+forced-choice and the betting paradigms would be an important next
+step.
 
@@
 Of course, our preliminary experiment should only serve as a start and more empirical data are needed for a robust pragmatic theory of people's reasoning about referential expressions.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "main"
+%%% TeX-PDF-mode: t
+%%% End:

### 2013-10-30 04_experiment.tex
Sec 4-6 proof

@@
 
-We recruited 1032 US Participants via Amazon Mechanical Turk, an online crowd-sourcing tool.
-Each participant played a one-shot referential game either as a speaker or as a listener and received a payment of 5 cents.
-If a participant contributed multiple trials, only the first was included.
-All participants passed the attention check described below.
+We recruited 1032 US participants via Amazon's Mechanical Turk, an
+online crowd-sourcing tool. Each participant played a one-shot
+referential game either as a speaker or as a listener and received a
+payment of 5 cents. If a participant contributed multiple trials, only
+the first was included. All participants passed the attention check
+described below.
 
@@
 
-Each participant saw a context consisting of three objects (Figure \ref{exp_context} as an example) and reported on the number of objects in the context as an attention check.
-Participants in the listener conditions were told to imagine someone was talking to them and used some utterance (depending on the exact condition) to refer to exactly one of the objects in the context.
-Then they were asked to choose what they thought to be the intended referent.
-Participants in the speaker conditions were told to imagine they were talking to someone and they wanted to refer to one of the objects (indicated by a small red arrow\footnote{Participants were told that the arrow was only for illustration and thus the person they were talking to could not see it.}).
-Then they were asked to choose between two words to refer to it.
+Each participant saw a context consisting of three objects (Figure
+\ref{exp_context} as an example) and reported on the number of objects
+in the context as an attention check. Participants in the listener
+conditions were told to imagine someone was talking to them who used
+some utterance (depending on the specific condition tested) to refer
+to exactly one of the objects in the context. Then they were asked to
+choose what they thought to be the intended referent. Participants in
+the speaker conditions were told to imagine they were talking to
+someone and they wanted to refer to one of the objects (indicated by a
+small red arrow\footnote{Participants were told that the arrow was
+  only for illustration and thus the person they were talking to could
+  not see it.}). Then they were asked to choose between two words to
+refer to it.
 
@@
 
-Each context had a square and a circle sharing the same color, and another circle having a different color.
-Hence each object had two features: shape (square/circle) and color (green/blue) and each context had two ambiguous (shared by two objects) features.
-In the listener salience condition, participants were told that the person talking to them was using a foreign language they could not understand, and in the other two listener conditions, the utterances used were the ambiguous feature of shape and color, respectively.
-In the speaker conditions, the target could be any of the three objects: the one with the unique color, the one with the unique shape and the one with both features shared.
-The two words for the participants to choose between were the features of color and shape of the target object.
+Each context had a square and a circle sharing the same color, and
+another circle having a different color. Hence each object had two
+features: shape (square/circle) and color (green/blue) and each
+context had two ambiguous (shared by two objects) features. In the
+listener salience condition, participants were told that the person
+talking to them was using a foreign language they could not
+understand, and in the other two listener conditions, the utterances
+used were the ambiguous feature of shape and color, respectively. In
+the speaker conditions, the target could be any of the three objects:
+the one with the unique color, the one with the unique shape and the
+one with both features shared. The two words for the participants to
+choose between were the features of color and shape of the target
+object.
 
-In order to minimize the effect of the confounding factors, we counterbalanced the positions and colors of the objects, as well as the orders of the candidate words in the speaker conditions.
-Thus the unique color in a context could be either blue (as in Fig.
-\ref{exp_context}) or green and the order of the objects could be any of the permutations\footnote{Due to the scale of our preliminary experiment we did not counterbalance the shape of the objects, i.e., a context always has one square and two circles.
-}.
+In order to minimize the effect of the confounding factors, we
+counterbalanced the positions and colors of the objects, as well as
+the orders of the candidate words in the speaker conditions. Thus the
+unique color in a context could be either blue (as in
+Fig. \ref{exp_context}) or green and the order of the objects could be
+any of the permutations.\footnote{Due to the scale of our preliminary
+  experiment we did not counterbalance the shape of the objects, i.e.,
+  a context always had one square and two circles.}
 
@@
 
-There are four major designs in our own experiment that differ from the original experiment in \cite{Frank}.
-First of all, the original experiment used the betting paradigm, i.e., participants were asked to bet over a range of options and they were instructed that the amount of money that is bet on an option should reflect their confidence that the option is correct.
-Even though the betting paradigm has the merit of providing us with graded reponses from each individual participant, the caveat is that it is unclear whether it measures the belief or the action.
-This can lead to confusion when we are to fit the model predictions to the empirical data without knowing whether they are directly comparable.
-In addition, since the betting paradigm is more or less introspective in nature, it tends to be not very accurate.
-Thus we used forced choice in our experiment instead, which clearly measures the action and does not rely on introspection.
-Secondly, since we decided to investigate the influence of the speaker's preference as well as the listener's perceptual salience, we focused on contexts equivalent to Figure \ref{context}, which are most typical in Gricean accounts of scalar implicature, and examined different features separately.
-Thirdly, we only included the features of color and shape to ensure the vocabulary to be common knowledge, excluding the feature of texture.
-Finally, we did not use dotted line surrounding an object as the way to indicate the target to the speaker, as it could emphasize the feature of shape and thus add confound in the experiment.
-Instead we used a small arrow pointing to the target.
+There are four major design choices in our own experiment that differ
+from the original experiment in \cite{Frank}. First of all, the
+original experiment used a \emph{betting paradigm}, i.e., participants
+were asked to bet over a range of options and they were instructed
+that the amount of money that is bet on an option should reflect their
+confidence that the option is correct. Even though the betting
+paradigm has the merit of providing us with graded reponses from each
+individual participant, the caveat is that it is unclear whether it
+measures beliefs or actions. This can lead to confusion when we are to
+fit the model predictions to the empirical data without knowing
+whether they are directly comparable. In addition, since the betting
+paradigm is more or less introspective in nature, it tends to be not
+very accurate. Thus we used a \emph{forced choice design} instead,
+which clearly measures the action, does not rely on introspection and,
+technically speaking, provides for a straightforward likelihood
+function to be used in subsequent statistical model comparison.
+Secondly, since we decided to investigate the influence of the
+speaker's preference as well as the listener's perceptual salience, we
+focused on contexts equivalent to Figure \ref{context}, which require
+reasoning that is highly reminiscent of Gricean accounts of scalar
+implicature calculation, and examined different features separately.
+Thirdly, we only included the features of color and shape to ensure
+the vocabulary to be common knowledge, excluding the feature of
+texture. Finally, we did not use dotted lines surrounding an object as
+the way to indicate the target to the speaker, as \cite{Frank} did,
+because that might unduly emphasize the feature of shape and thus be a
+confound in the experiment. Instead we used a small arrow pointing to
+the target.
 
-Despite these differences, the remaining aspects in our experiment were almost identical to those of the original one, e.g. the phrasing of the instructions.
+Despite these differences, the remaining aspects in our experiment
+were almost identical to those of the original one, e.g. the phrasing
+of the instructions.
 
-\
-
-The results of the speaker and listener conditions of our experiment are shown as follows.
+The results of the speaker and listener conditions of our experiment
+are as follows.
 
@@
      \hline
-\textcolor{green!65}{\Large{$\blacksquare$}}    & $9$        &   $135$   & -- & -- &144\\
+\textcolor{green!65}{\large{$\blacksquare$}}    & $9$        &   $135$   & -- & -- &144\\
 \textcolor{blue!65}{\Huge{$\bullet$}}           & --        &   --      & $119$ & $25$ & 144\\
@@
 
-From the data in Table \ref{table:speaker}, we can see that the speakers tend to choose the optimal feature more often when the target has the unique shape than when it has the unique color ($\chi^2=8.5, p<.01$).
-Even though they seem to prefer the feature of shape when the target's both features are shared, the difference is not statistically significant from uniform random choice ($\chi^2=2.25, p=.13$).
+From the data in Table \ref{table:speaker}, we can see that the
+speakers tend to choose the optimal feature more often when the target
+has the unique shape than when it has the unique color ($\chi^2=8.5,
+p<.01$). Even though they seem to prefer the feature of shape when
+both of the target's features are shared, the difference is not
+statistically significant from uniform random choice ($\chi^2=2.25,
+p=.13$).
 
@@
  Salience   \quad  & \quad  $71$      \quad    &   $30$    \quad     & $139$ \quad     & $240$      \\
- ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & $0$ \quad     & $180$     \\
+ ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & -- \quad     & $180$     \\
  ``Circle''   &  \quad $1$        &    $62$        & $117$ \quad     & $180$
@@
 
-The result of the salience condition will be used as the empirical estimation of contextual salience.
-For the other two conditions, the object with both features shared (the green circle in the above table, but note that we counterbalanced colors) is what the Gricean pragmatic account predicts to be the target.
-The result in Table \ref{table:listener} shows that when the message is ``Green'', listeners prefer the green circle which is predicted by the Gricean pragmatics, while they tend to stick to the blue circle which is more perceptually salient when the message is ``Circle''.
-The behavioral patterns in both conditions are significantly different from uniform random choice, and they significantly differ from each other in respect of whether they conform to the predictions by Gricean pragmatic theory.
+The result of the salience condition will be used as the empirical
+estimation of contextual salience. For the other two conditions, the
+object with both features shared (the green circle in the above table,
+but note that we counterbalanced colors) is what the Gricean pragmatic
+account predicts to be the target. The result in Table
+\ref{table:listener} shows that when the message is ``Green'',
+listeners prefer the green circle which is predicted by the Gricean
+pragmatics, while they tend to stick to the blue circle which is more
+perceptually salient when the message is ``Circle''. The behavioral
+patterns in both conditions are significantly different from uniform
+random choice, and they significantly differ from each other in
+whether they conform to the predictions by Gricean pragmatic theory.
 

### 2013-10-30 06_conclusion.tex
Sec 4-6 proof

--- 2013-10-14 06_conclusion.tex
@@
-Our experiment and model comparison suggest the following: (1) On the conceptual level, 
-it is helpful to clarify the distinction between belief, goal and action, and the Bayesian framework provides us with a natural way to structure these components into a formal model.
-By examining each component systematically, we can explicitly spell out the underlying assumptions and formulate alternative hypotheses to be tested.
-(2) On the empirical level, we tested the RSA model with a family of variants motivated from the game-theoretic pragmatic perspectives on communication, and in particular, we investigated the speaker's preference as well as the listener's perceptual salience.
-Our data showed a more intricate picture about the roles various factors play in pragmatic reasoning about referential expressions --- what is commonly known is the Gricean pragmatic components, while either side have his own private bias unknown to the other.
-Also, an action-based view might better reflect the goal of communication, at least in the forced-choice experimental setting.
+Our experiment and model comparison suggest the following: (1) On the
+conceptual level, it is helpful to clarify the distinction between
+belief, goal and action, and the Bayesian framework provides us with a
+natural way to structure these components into a formal model. By
+examining each component systematically, we can explicitly spell out
+the underlying assumptions and formulate alternative hypotheses to be
+tested. (2) On the empirical level, we tested the RSA model with a
+family of variants motivated from a game theoretic perspective on
+communication, and in particular, we investigated the speaker's
+preference as well as the listener's perceptual salience. Our data
+showed a more intricate picture of the roles that various factors play
+in pragmatic reasoning about referential expressions. We found
+evidence for a correlation between the speaker's and the listener's
+rationality. Either side appears to have his own biases, but appears
+to be negligent of the other's. Also, an action-based view might
+better reflect the goal of communication, at least in the
+forced-choice experimental setting. Understanding the relation between
+the forced-choice and the betting paradigms would be an important next
+step to gain more insight into the differences between action- and
+belief-based goal structures.
 
-Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to a better understanding of the phenomenon and further improvement in the formal model's predictive power.
-Of course, more empirical data are needed for a robust pragmatic theory of people's reasoning about referential expressions.
+Our work shows that careful conceptual analysis of the design choices
+for quantitative models can lead to a better understanding of the
+phenomenon and further improvement in the formal model's predictive
+power. Of course, our restricted empirical data can only serve as a
+start and more data is needed to fuel further model comparison towards
+a more robust pragmatic theory of people's probabilistic reasoning
+about referential expressions.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "main"
+%%% TeX-PDF-mode: t
+%%% End:

### 2014-03-14 02_RSA.tex
minor corrections
several footnotes for reviewers' comments

@@
 denoting a feature, i.e., color or shape, of the target) to the
-listener. The listener, who has does not know which object is the
+listener. The listener, who does not know which object is the
 target, needs to recover the target based on the speaker's choice of
@@
 degree of rationality, i.e., to what extent the speaker sticks to the
-strict optimum. The cost term is used to encode preference in
+strict optimum\footnote{Mathematically, it can be shown that when $\lambda_\mathrm{S}=0$, $\sigma(m \mid t)$ is uniformly distributed over all messages, meaning that the speaker just randomly selects one of the messages without any optimality consideration.
+On the other hand, when $\lambda_\mathrm{S}\rightarrow \infty$, 
+$\sigma(m \mid t)$ is non-zero only when $m$ yields the maximal utility, 
+meaning that the speaker only selects the strictly optimal option(s).
+This rule is widely used in the AI literature (e.g., \cite{Sutton1998RL}) and 
+is also known as the \emph{exponential rule}.}.
+The cost term is used to encode preference in
 different utterances, be it about the utterances' lengths or syntactic
@@
 
-From (\ref{Bayesian_literal})-(\ref{Bayesian_sender_softmax}) we obtain the speaker's production rule:
+From (\ref{Bayesian_literal})-(\ref{Bayesian_sender_softmax}) we obtain the speaker's production rule\footnote{Note that we can further simplify (\ref{Bayesian_speaker}) as follows: $\exp(\lambda_\mathrm{S} \cdot (\log \mathcal{U}(t\mid \intp{m})-\mbox{Cost}(m)))
+=\exp(\lambda_\mathrm{S} \log \mathcal{U}(t\mid \intp{m}) \cdot \exp(-\lambda_\mathrm{S}\cdot \mbox{Cost}(m))
+=(\mathcal{U}(t\mid \intp{m}))^{\lambda_\mathrm{S}}\cdot \exp(-\lambda_\mathrm{S}\cdot \mbox{Cost}(m))$.
+Incidentally, if we do not take costs into account, i.e., let $\mbox{Cost}(m)\equiv 0$, then (\ref{Bayesian_speaker}) reduces to $\sigma(m \mid t) \propto (\mathcal{U}(t\mid \intp{m}))^{\lambda_\mathrm{S}}$, which is effectively a \emph{power rule} that soft-maximizes $\mathcal{U}(t\mid \intp{m})$.
+In this respect
+(\ref{Bayesian_speaker}) can be seen as an action-oriented speaker model (the exact definition of which will be given in the next section) equipped with 
+a different kind of soft-max rule. However, we note that this is a mere technical coincidence, and as a first step, throughout this paper we only consider the exponential 
+rule as the soft-max rule. \cite{Worthy2008ratio} provides a detailed discussion and empirical analyses on the differences between the exponential and power rules.}:
 

### 2014-03-14 03_alternatives.tex
minor corrections
several footnotes for reviewers' comments

@@
 \cite{FrankeJager2013:Pragmatic-Back-} for overview and comparison
-of game theoretic and Bayesian models.} Doing so
-let's us describe a class of potential alternatives to RSA. We will
-then compare the predictive power of these given relevant empirical
-data in later sections.
+of game theoretic and Bayesian models.} We will describe a class of potential 
+alternatives to RSA and then compare the predictive power of these alternatives 
+on relevant empirical data in later sections.
 
@@
 since neither his beliefs nor actions depend on it. A natural variant,
-which is used in some game theoretic models, would replace the literal
+which is used in some game theoretic models, would be to replace the literal
 listener's uniform distribution $\mathcal{U}(t)$ in
@@
 the probability of the listener choosing the intended
-referent. Formally speaking, we have the action-oriented speaker model
+referent.
+Formally speaking, we have the action-oriented speaker model\footnote{There might be concern about whether there should be another soft-max step for the literal listener's choice probability $\rho_0(t\mid m)$ in order for it to be action-oriented.
+This depends on the precise conceptual status of the literal listener in the model and to our knowledge there is no consensus yet.
+We tend to think of a literal listener as an automatic subconscious process that involves no optimality reasoning or decision-making.
+Thus we consider it appropriate to use $\rho_0(t\mid m)$ as the action of the literal listener.
+However, we do agree that in principle there can be a model whose literal listener does a soft-max over the salience prior, and such a model can be similarly included in the model comparison.}:
 
@@
 Hence we have four types of speaker models that differ in either the speaker's belief about the literal listener, or the speaker's goal of communication.
-We now introduce a uniform notation $\sigma_{xy},\ x\in\{a,b\},y\in\{\mathcal{U},\mathcal{S}\}$ for them:
+We now introduce a uniform notation $\sigma_{xy},\ x\in\{\mathrm{a},\mathrm{b}\},y\in\{\mathcal{U},\mathcal{S}\}$ for them:
 \begin{equation} \label{Bayesian_speaker_uniform_action}

### 2014-03-14 02_RSA.tex
thanks, references, typos

@@
 such as informativeness and how different pieces of information should
-be integrated.
+be integrated.\todo{discuss Frank and Goodman's experiments and
+  results (in as far possible) here, before venturing into the model}
 
@@
 This rule is widely used in the AI literature (e.g., \cite{Sutton1998RL}) and 
-is also known as the \emph{exponential rule}.}.
+is also known as \emph{exponential choice rule}.}.
 The cost term is used to encode preference in
@@
 
-From (\ref{Bayesian_literal})-(\ref{Bayesian_sender_softmax}) we obtain the speaker's production rule\footnote{Note that we can further simplify (\ref{Bayesian_speaker}) as follows: $\exp(\lambda_\mathrm{S} \cdot (\log \mathcal{U}(t\mid \intp{m})-\mbox{Cost}(m)))
-=\exp(\lambda_\mathrm{S} \log \mathcal{U}(t\mid \intp{m}) \cdot \exp(-\lambda_\mathrm{S}\cdot \mbox{Cost}(m))
-=(\mathcal{U}(t\mid \intp{m}))^{\lambda_\mathrm{S}}\cdot \exp(-\lambda_\mathrm{S}\cdot \mbox{Cost}(m))$.
-Incidentally, if we do not take costs into account, i.e., let $\mbox{Cost}(m)\equiv 0$, then (\ref{Bayesian_speaker}) reduces to $\sigma(m \mid t) \propto (\mathcal{U}(t\mid \intp{m}))^{\lambda_\mathrm{S}}$, which is effectively a \emph{power rule} that soft-maximizes $\mathcal{U}(t\mid \intp{m})$.
-In this respect
-(\ref{Bayesian_speaker}) can be seen as an action-oriented speaker model (the exact definition of which will be given in the next section) equipped with 
-a different kind of soft-max rule. However, we note that this is a mere technical coincidence, and as a first step, throughout this paper we only consider the exponential 
-rule as the soft-max rule. \cite{Worthy2008ratio} provides a detailed discussion and empirical analyses on the differences between the exponential and power rules.}:
+From (\ref{Bayesian_literal})-(\ref{Bayesian_sender_softmax}) we
+obtain the speaker's production rule:\footnote{Note that we can further
+  simplify (\ref{Bayesian_speaker}) as follows:
+  $\exp(\lambda_\mathrm{S} \cdot (\log \mathcal{U}(t\mid
+  \intp{m})-\mbox{Cost}(m))) =\exp(\lambda_\mathrm{S} \log
+  \mathcal{U}(t\mid \intp{m}) \cdot \exp(-\lambda_\mathrm{S}\cdot
+  \mbox{Cost}(m)) =(\mathcal{U}(t\mid
+  \intp{m}))^{\lambda_\mathrm{S}}\cdot \exp(-\lambda_\mathrm{S}\cdot
+  \mbox{Cost}(m))$. Incidentally, if we do not take costs into
+  account, i.e., let $\mbox{Cost}(m) = 0$, then
+  (\ref{Bayesian_speaker}) reduces to $\sigma(m \mid t) \propto
+  (\mathcal{U}(t\mid \intp{m}))^{\lambda_\mathrm{S}}$, which is
+  effectively a \emph{power rule} that soft-maximizes
+  $\mathcal{U}(t\mid \intp{m})$. In this respect
+  (\ref{Bayesian_speaker}) can be seen as an action-oriented speaker
+  model (the exact definition of which will be given in the next
+  section) equipped with a different kind of soft-max rule. However,
+  we note that this is a mere technical coincidence, and as a first
+  step, throughout this paper we only consider the exponential rule as
+  the soft-max rule. \cite{Worthy2008ratio} provides a detailed
+  discussion and empirical analyses on the differences between the
+  exponential and power rules.}
 

### 2014-03-14 main.tex
minor corrections
several footnotes for reviewers' comments
thanks, references, typos

@@
 \title{Variations on a Bayesian Theme: Comparing Bayesian Models of
-  Referential Reasoning}
-\titlerunning{Variation on a Bayesian Theme}
-\author{Ciyang Qing \and Michael Franke\thanks{Thanks to \dots % Judith Degen,
-  % Noah Goodman, Daniel Lassiter [?], audience of the workshop, Henk
-  % & Hans-Christian for organizing, reviewers \dots 
-}}
-\institute{Institute for Logic, Language and Computation\\Universtity
-  of Amsterdam, The Netherlands}
+  Referential Reasoning} \titlerunning{Variation on a Bayesian Theme}
+\author{Ciyang Qing \and Michael Franke\thanks{We are indebted to
+    Judith Degen, Michael C.~Frank, Noah D.~Goodman and Daniel
+    Lassiter, two anonymous reviewers, and the audience of the
+    ESSLLI workshop ``Bayesian Natural Language Semantics and
+    Pragmatics'' for stimulating feedback and discussion. Many thanks
+    also to Henk Zeevat and Hans-Christian for organizing mentioned
+    workshop, and to Will Frager for help accessing Amazon Mechanical
+    Turk. Michael Franke gratefully acknowledges financial support by
+    NWO-VENI grant 275-80-004.}}  \institute{Institute for Logic, Language and
+  Computation\\Universtity of Amsterdam, The Netherlands}
 

### 2014-03-15 02_RSA.tex
intro to FG's experiment

@@
 the rational expectation of the likelihood of the speaker making the
-utterance for each object. The latter line of reasoning is crucial to
-the classic Gricean account of scalar implicatures, and the major
-challenges in terms of formal modeling are how to quantify notions
+utterance for each object.
+
+The experiment in \cite{Frank} confirmed the above intuitions about the 
+referential games. 
+In their study, 745 US participants were recruited via 
+Amazon’s Mechanical Turk to take part in a one-shot referential game, either 
+as the speaker or the listener. 
+The context of the game always consisted of three objects. 
+Each object had three features: color (red, blue, or green), shape 
+(circle, square, or cloud), and texture (solid, or polka-dot, or striped).
+Every context had one feature fixed for all objects and the other two features
+systematically varied. 
+The objects were randomly permuted and the target object was randomly assigned. 
+Depending on how many objects share the same value of a feature with the target, 
+the contexts can be classified into 7 categories: 1/1, 1/2 (the target's 
+first feature is unique, and there is another object sharing the same value of
+the second feature with the target), 1/3, 2/2 with the same object sharing both
+features with the target (which means there is another object identical to the 
+target), 2/2 with different objects sharing the features (such as in 
+Fig.~\ref{context}), 2/3 and 3/3. 
+The speakers were asked to choose a feature to refer to the target object. 
+In order to measure the perceptual salience, some of the listeners were told 
+that the speaker was using a language that they did not understand and then 
+they were asked to guess the intended referent. 
+Other listeners were told a feature of the target and asked to choose the 
+intended referent.
+
+The result of the experiment showed that speakers generally followed 
+Grice's Maxim of Quantity, i.e., they tried to be informative by choosing the 
+most unique feature of the target, and the listeners were influenced by both 
+the perceptual salience and the informativeness of the utterance that the 
+speaker produced. 
+However, the data, being quantitative in nature, are also full of apparent 
+exceptions. 
+This means that the empirical data cannot be fully explained by the classic 
+Gricean account of scalar implicatures, and the major challenges in terms of 
+formal modeling include how to quantify notions
 such as informativeness and how different pieces of information should
-be integrated.\todo{discuss Frank and Goodman's experiments and
-  results (in as far possible) here, before venturing into the model}
+be integrated.
 

### 2014-03-17 02_RSA.tex
added conclusions, minor changes here and there

--- 2014-03-14 02_RSA.tex
@@
 Referential games are confined interactive reasoning tasks used to
-study pragmatic inference in a controlled environment \cite{Frank}. A
-referential game consists of a set of objects called a
-\textit{context}. A context contains different shapes of different
-colors in various arrangements, such as shown in Figure~\ref{context}.
-The speaker in the game is asked to refer to one of the objects (the
-\emph{target}) by uttering an expression (typically a single word
-denoting a feature, i.e., color or shape, of the target) to the
-listener. The listener, who does not know which object is the
-target, needs to recover the target based on the speaker's choice of
-expression.
+study pragmatic inference in a controlled environment
+\cite{StillerGoodman2011:Ad-Hoc-Scalar-I,Frank,Degen}. A referential game
+consists of a set of objects called a \textit{context}. For example, a
+context might contain different shapes of different colors in various
+arrangements, such as shown in Figure~\ref{context}.  The speaker in
+the game is asked to refer to one of the objects (the \emph{target})
+by uttering an expression (typically a single word denoting a feature,
+i.e., color or shape, of the target) to the listener. The listener,
+who does not know which object is the target, needs to recover the
+target based on the speaker's choice of expression.
 
@@
   }
-  \caption{A simple referential game}
+  \caption{A simple referential game, as used in \cite{Frank}}
 \end{figure}
@@
 the rational expectation of the likelihood of the speaker making the
-utterance for each object. The latter line of reasoning is crucial to
-the classic Gricean account of scalar implicatures, and the major
-challenges in terms of formal modeling are how to quantify notions
-such as informativeness and how different pieces of information should
-be integrated.\todo{discuss Frank and Goodman's experiments and
-  results (in as far possible) here, before venturing into the model}
+utterance for each object.
 
-The RSA model addresses the problems by using information-theoretic
-concepts to measure informativeness and Bayesian inference to
-integrate different sources of information
+To test these intuitions, Frank \& Goodman recruited 745 US
+participants via Amazon's Mechanical Turk to take part in a one-shot
+referential game, either as the speaker or the listener \cite{Frank}.
+The context of the game always consisted of three objects (e.g.,
+Figure~\ref{context}).  Each object had three features: color (red,
+blue, or green), shape (circle, square, or cloud), and texture (solid,
+or polka-dot, or striped).  Every context had one feature fixed for
+all objects and the other two features systematically varied.  The
+objects were randomly permuted and the target object was randomly
+assigned.  Depending on how many objects shared the same value of a
+feature with the target, contexts were classified into 7 categories:
+1/1, 1/2 (the target's first feature is unique, and there is another
+object sharing the same value of the second feature with the target),
+1/3, 2/2 with the same object sharing both features with the target
+(which means there is another object identical to the target), 2/2
+with different objects sharing the features (such as in
+Fig.~\ref{context}), 2/3 and 3/3. The speakers were asked to choose a
+feature to refer to the target object.\footnote{More accurately,
+  participants were ask to \emph{bet} on features and referents. We
+  will discuss this design choice when contrasting it with our own
+  replication experiment in Section~\ref{sec:experiment}.}  In order
+to measure the perceptual salience, some of the listeners were told
+that the speaker was using a language that they did not understand and
+then they were asked to guess the intended referent.  Other listeners
+were told a feature of the target and asked to choose the intended
+referent.
+
+The result of the experiment showed that speakers generally followed
+Grice's Maxim of Quantity, i.e., they tried to be informative by
+choosing the most unique feature of the target, and that listeners
+were influenced by both the perceptual salience and the
+informativeness of the utterance that the speaker produced. However,
+the data, being quantitative in nature, cannot be fully explained by a
+classical qualitative Gricean account of scalar implicatures, so that
+a major challenge in terms of formal modeling becomes how to quantify
+notions such as informativeness and perceptual salience, and how these
+should be integrated to yield qualitative predictions.
+
+The rational speech-act model (RSA) model addresses these issues by
+using information-theoretic concepts to measure informativeness and
+Bayesian inference to integrate different sources of information
 \cite{Frank2009,Frank,Bergen2012,GoodmanStuhlmuller2013:Knowledge-and-I}. In
@@
 degree of rationality, i.e., to what extent the speaker sticks to the
-strict optimum\footnote{Mathematically, it can be shown that when $\lambda_\mathrm{S}=0$, $\sigma(m \mid t)$ is uniformly distributed over all messages, meaning that the speaker just randomly selects one of the messages without any optimality consideration.
+strict optimum.\footnote{Mathematically, it can be shown that when $\lambda_\mathrm{S}=0$, $\sigma(m \mid t)$ is uniformly distributed over all messages, meaning that the speaker just randomly selects one of the messages without any optimality consideration.
 On the other hand, when $\lambda_\mathrm{S}\rightarrow \infty$, 
@@
 This rule is widely used in the AI literature (e.g., \cite{Sutton1998RL}) and 
-is also known as \emph{exponential choice rule}.}.
+is also known as \emph{exponential choice rule}.}
 The cost term is used to encode preference in

### 2014-03-17 03_alternatives.tex
added conclusions, minor changes here and there

@@
 the probability of the listener choosing the intended
-referent.
-Formally speaking, we have the action-oriented speaker model\footnote{There might be concern about whether there should be another soft-max step for the literal listener's choice probability $\rho_0(t\mid m)$ in order for it to be action-oriented.
-This depends on the precise conceptual status of the literal listener in the model and to our knowledge there is no consensus yet.
-We tend to think of a literal listener as an automatic subconscious process that involves no optimality reasoning or decision-making.
-Thus we consider it appropriate to use $\rho_0(t\mid m)$ as the action of the literal listener.
-However, we do agree that in principle there can be a model whose literal listener does a soft-max over the salience prior, and such a model can be similarly included in the model comparison.}:
+referent. Formally speaking, we have the action-oriented speaker
+model:\footnote{There might be concern about whether there should be
+  another soft-max step for the literal listener's choice probability
+  $\rho_0(t\mid m)$ in order for it to be action-oriented.  This
+  depends on the precise conceptual status of the literal listener in
+  the model and to our knowledge there is no consensus yet. We tend to
+  think of a literal listener as an automatic subconscious process
+  that involves no optimality reasoning or decision-making. Thus we
+  consider it appropriate to use $\rho_0(t\mid m)$ as the action of
+  the literal listener. However, we do agree that in principle there
+  can be a model whose literal listener does a soft-max over the
+  salience prior, and such a model can be similarly included in the
+  model comparison.}
 

### 2014-03-17 04_experiment.tex
minor corrections
added conclusions, minor changes here and there

@@
-In the previous section we analyzed the design of the original RSA model, spelled out the relation and distinction between belief, goal and action, and proposed a familiy of alternative models based on different interpretation of these notions.
-This gives rise to a series of interesting empirical questions regarding the underlying assumptions of the models.
-In order to gain insight into these questions by comparing the predictive power of different models, we conducted the following experiment to collect empirical data, which we will use to test the predictions of different models in the next section.
+In the previous section we analyzed the design of the original RSA
+model, spelled out the relation and distinction between belief, goal
+and action, and proposed a familiy of alternative models based on
+different interpretation of these notions. This gives rise to a series
+of interesting empirical questions regarding the underlying
+assumptions of the models. In order to gain insight into these
+questions by comparing the predictive power of different models, we
+conducted the following experiment to collect empirical data, which we
+will use to test the predictions of different models in the next
+section.
 
@@
 
-We recruited 1032 US participants via Amazon's Mechanical Turk, an
-online crowd-sourcing tool. Each participant played a one-shot
-referential game either as a speaker or as a listener and received a
-payment of 5 cents. If a participant contributed multiple trials, only
-the first was included. All participants passed the attention check
-described below.
+We recruited 1032 US participants via Amazon's Mechanical Turk. Each
+participant played a one-shot referential game either as a speaker or
+as a listener and received a payment of 5 cents. If a participant
+contributed multiple trials, only the first was included. All
+participants passed the attention check described below.
 
@@
 texture. Finally, we did not use dotted lines surrounding an object as
-the way to indicate the target to the speaker, as \cite{Frank} did,
+the way to indicate the target to the speaker, as done in \cite{Frank},
 because that might unduly emphasize the feature of shape and thus be a
@@
  Salience   \quad  & \quad  $71$      \quad    &   $30$    \quad     & $139$ \quad     & $240$      \\
- ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & -- \quad     & $180$     \\
+ ``Green''   \quad  & \quad  $65$      \quad    &   $115$    \quad     & $0$ \quad     & $180$     \\
  ``Circle''   &  \quad $1$        &    $62$        & $117$ \quad     & $180$

### 2014-03-17 06_conclusion.tex
added conclusions, minor changes here and there

@@
 
-Our work shows that careful conceptual analysis of the design choices
-for quantitative models can lead to a better understanding of the
-phenomenon and further improvement in the formal model's predictive
-power. Of course, our restricted empirical data can only serve as a
-start and more data is needed to fuel further model comparison towards
-a more robust pragmatic theory of people's probabilistic reasoning
-about referential expressions.
+The set of models we were able to compare explicitly here does clearly
+not exhaust the space of plausible candidates. For example, we
+excluded listener models in which the literal listener takes the
+salience prior into account. As an anonymous reviewer points out, this
+may seem paradoxical because the best listener models of our
+comparison do assume that the actual pragmatic listener takes salience
+into account, but does not believe that the speaker believes that he
+does. But actually, there is no friction here at all, because the
+listener model that is embedded in a speaker model is a different one
+from the full pragmatic listener model. If we think of the pragmatic
+listener as simulating his own hypothetical speaker behavior, there is
+nothing puzzling about factoring in salience only at a higher level of
+pragmatic inference.
+
+A similar issue arises with respect to the rationality parameters in
+action-based listener models. We assumed that there are two
+rationality parameters, one for the listener's actual rationality and
+one for the listener's beliefs about the speaker's rationality. But we
+also saw that a model that correlates the set of joint values of these
+parameter pairs gave the best predictions of our data. This raises
+many interesting issues for further research, most of which we cannot
+answer here. The most obvious idea is, of course, to equate the
+speaker's and the listener's lambda and so that the listener would
+believe that the speaker is exactly as rational as the listener
+himself. Surprisingly, this model achieves a very poor fit on our
+data, despite its parsimony. More research into correlations in
+perspective-taking and beliefs in rationality will hopefully shed more
+light on this fascinating issue.
+
+In summary, our work shows that careful conceptual analysis of the
+design choices for quantitative models can lead to a better
+understanding of the phenomenon and further improvement in the formal
+model's predictive power. Of course, our restricted empirical data can
+only serve as a start and more data is needed to fuel further model
+comparison towards a more robust pragmatic theory of people's
+probabilistic reasoning about referential expressions.
 

### 2014-03-17 06_conclusion.tex
added conclusions, minor changes here and there

@@
-Our experiment and model comparison suggest the following: (1) On the
+Our experiment and model comparison suggest the following. (1) On the
 conceptual level, it is helpful to clarify the distinction between
@@
 better reflect the goal of communication, at least in the
-forced-choice experimental setting. Understanding the relation between
-the forced-choice and the betting paradigms would be an important next
-step to gain more insight into the differences between action- and
-belief-based goal structures.
+forced-choice experimental setting that we used. Understanding the
+relation between the forced-choice and the betting paradigms would be
+an important next step to gain more insight into the differences
+between action- and belief-based goal structures.
 

### 2014-03-18 02_RSA.tex
experiment-related revision

@@
 
+\begin{figure}[h] \label{fig:fg-all}
+  \centering \subfigure[1/1]{\label{fg-1-1}
+  \includegraphics[scale=0.2]{pics/fg-1-1.jpg}
+} \subfigure[1/2]{\label{fg-1-2}
+  \includegraphics[scale=0.2]{pics/fg-1-2.jpg}
+} \subfigure[1/3]{\label{fg-1-3}
+  \includegraphics[scale=0.2]{pics/fg-1-3.jpg}
+}
+
+\subfigure[2/2: identical]{\label{fg-2-2-dup}
+  \includegraphics[scale=0.2]{pics/fg-2-2-dup.jpg}
+} \subfigure[2/2: different]{\label{fg-2-2-crit}
+  \includegraphics[scale=0.2]{pics/fg-2-2-crit.jpg}
+} \subfigure[2/3]{\label{fg-2-3}
+  \includegraphics[scale=0.2]{pics/fg-2-3.jpg}
+}
+
+\subfigure[3/3]{\label{fg-3-3}
+  \includegraphics[scale=0.2]{pics/fg-3-3.jpg}
+}
+
+  \caption{Sample contexts for the 7 conditions used in \cite{Frank}. 
+  The targets are indicated by dotted lines. 
+  Condition $n/m$ means that the target has one feature shared 
+  by $n$ objects and another feature shared by $m$ objects in the context
+  (including the target itself).
+  For $2/2$, either there is an object identical to the target, or the features are shared by different objects.}
+\end{figure}
+
 To test these intuitions, Frank \& Goodman recruited 745 US
@@
 assigned.  Depending on how many objects shared the same value of a
-feature with the target, contexts were classified into 7 categories:
+feature with the target, contexts were classified into 7 categories
+(Fig.~\ref{fig:fg-all}):
 1/1, 1/2 (the target's first feature is unique, and there is another
@@
 (which means there is another object identical to the target), 2/2
-with different objects sharing the features (such as in
-Fig.~\ref{context}), 2/3 and 3/3. The speakers were asked to choose a
+with different objects sharing the features, 2/3 and 3/3.  
+The speakers were asked to choose a
 feature to refer to the target object.\footnote{More accurately,
   participants were ask to \emph{bet} on features and referents. We
-  will discuss this design choice when contrasting it with our own
-  replication experiment in Section~\ref{sec:experiment}.}  In order
+  will discuss this design choice when contrasting it with our 
+  replication in Section~\ref{sec:experiment}.}  In order
 to measure the perceptual salience, some of the listeners were told
@@
 referent. 
+Note that listeners were presented with pictures without dotted lines 
+surrounding the target so they did not know the target.
+For each category in each of the speaker, salience and listener conditions, 
+there were roughly 50 responses\footnote{As we can see, despite the fairly 
+large number of participants recruited, the data in each condition 
+are not that many. Also there might be concern that some conditions might 
+look less interesting than others. Again we will discuss these issues when 
+we introduce our own replication in Section~\ref{sec:experiment}. Here
+we note that variety is important in order to test the general validity of 
+a pragmatic theory. 
+Thus the design used in \cite{Frank} served its purposes well, but may be 
+adapted when the goal is different.}.
 
@@
 Grice's Maxim of Quantity, i.e., they tried to be informative by
-choosing the most unique feature of the target, and that listeners
+choosing the most unique feature of the target\footnote{Speakers' preference
+was not taken into account in \cite{Frank}, so all features were randomly 
+assigned and later lumped together in the analysis.}, and that listeners
 were influenced by both the perceptual salience and the

### 2014-03-18 04_experiment.tex
experiment-related revision

@@
 implicature calculation, and examined different features separately.
-Thirdly, we only included the features of color and shape to ensure
-the vocabulary to be common knowledge, excluding the feature of
-texture. Finally, we did not use dotted lines surrounding an object as
+Thirdly, we only included the features of color and shape, with binary
+values (green/blue and square/circle) for two reasons. On the one hand, 
+we wanted to ensure the vocabulary to be common knowledge, so we excluded 
+the feature of texture. On the other hand, we excluded feature values red 
+and cloud as they are significantly more salient than the others, which 
+might override the pragmatic effect. 
+Finally, we did not use dotted lines surrounding an object as
 the way to indicate the target to the speaker, as done in \cite{Frank},
@@
 The results of the speaker and listener conditions of our experiment
-are as follows.
+are shown as follows.
 

### 2014-03-18 02_RSA.tex
final

@@
 feature with the target, contexts were classified into 7 categories
-(Fig.~\ref{fig:fg-all}):
-1/1, 1/2 (the target's first feature is unique, and there is another
-object sharing the same value of the second feature with the target),
-1/3, 2/2 with the same object sharing both features with the target
-(which means there is another object identical to the target), 2/2
-with different objects sharing the features, 2/3 and 3/3.  
-The speakers were asked to choose a
-feature to refer to the target object.\footnote{More accurately,
-  participants were ask to \emph{bet} on features and referents. We
-  will discuss this design choice when contrasting it with our 
-  replication in Section~\ref{sec:experiment}.}  In order
-to measure the perceptual salience, some of the listeners were told
-that the speaker was using a language that they did not understand and
-then they were asked to guess the intended referent.  Other listeners
-were told a feature of the target and asked to choose the intended
-referent. 
-Note that listeners were presented with pictures without dotted lines 
-surrounding the target so they did not know the target.
-For each category in each of the speaker, salience and listener conditions, 
-there were roughly 50 responses\footnote{As we can see, despite the fairly 
-large number of participants recruited, the data in each condition 
-are not that many. Also there might be concern that some conditions might 
-look less interesting than others. Again we will discuss these issues when 
-we introduce our own replication in Section~\ref{sec:experiment}. Here
-we note that variety is important in order to test the general validity of 
-a pragmatic theory. 
-Thus the design used in \cite{Frank} served its purposes well, but may be 
-adapted when the goal is different.}.
+(see Fig.~\ref{fig:fg-all}): 1/1, 1/2 (the target's first feature is
+unique, and there is another object sharing the same value of the
+second feature with the target), 1/3, 2/2 with the same object sharing
+both features with the target (which means there is another object
+identical to the target), 2/2 with different objects sharing the
+features, 2/3 and 3/3.  The speakers were asked to choose a feature to
+refer to the target object.\footnote{More accurately, participants
+  were ask to \emph{bet} on features and referents. We will discuss
+  this design choice when contrasting it with our replication in
+  Section~\ref{sec:experiment}.}  In order to measure the perceptual
+salience, some of the listeners were told that the speaker was using a
+language that they did not understand and then they were asked to
+guess the intended referent.  Other listeners were told a feature of
+the target and asked to choose the intended referent.  Note that
+listeners were presented with pictures without dotted lines
+surrounding the target so they did not know the target.  For each
+category in each of the speaker, salience and listener conditions,
+there were roughly 50 responses.\footnote{As we can see, despite the
+  fairly large number of participants recruited, the data in each
+  condition are not that many. Also there might be a concern that some
+  conditions might look less interesting than others. Again we will
+  discuss these issues when we introduce our own replication in
+  Section~\ref{sec:experiment}. Here we note that variety is important
+  in order to test the general validity of a pragmatic theory.  Thus
+  the design used in \cite{Frank} served its purposes well, but may be
+  adapted when the goal is different.}
 
@@
 Grice's Maxim of Quantity, i.e., they tried to be informative by
-choosing the most unique feature of the target\footnote{Speakers' preference
-was not taken into account in \cite{Frank}, so all features were randomly 
-assigned and later lumped together in the analysis.}, and that listeners
-were influenced by both the perceptual salience and the
-informativeness of the utterance that the speaker produced. However,
-the data, being quantitative in nature, cannot be fully explained by a
-classical qualitative Gricean account of scalar implicatures, so that
-a major challenge in terms of formal modeling becomes how to quantify
-notions such as informativeness and perceptual salience, and how these
-should be integrated to yield qualitative predictions.
+choosing the most unique feature of the target,\footnote{Speakers'
+  preference was not taken into account in \cite{Frank}, so all
+  features were randomly assigned and later lumped together in the
+  analysis.} and that listeners were influenced by both the perceptual
+salience and the informativeness of the utterance that the speaker
+produced. However, the data, being quantitative in nature, cannot be
+fully explained by a classical qualitative Gricean account of scalar
+implicatures, so that a major challenge in terms of formal modeling
+becomes how to quantify notions such as informativeness and perceptual
+salience, and how these should be integrated to yield qualitative
+predictions.
 

### 2014-03-18 response_to_reviewers.txt
response letter

Reviewer A:

> There is a large literature on perspective-taking and simulation in theory of 
> mind tasks, which would be informative to discuss in the context of this issue.

Indeed, it would have been great to discuss more on this, but since we are approaching the page limit we could not elaborate much further on this.

> First, the authors show that listeners take into account salience and that
> speakers do not model the listeners' salience sensitivity; but they do
> not evaluate whether listeners believe that speakers model their
> salience sensitivity (the relevant model comparisons are excluded from
> the analyses presented in the paper). It would be rather
> surprising if the speakers' model of the listeners was deficient in
> this respect, and the listeners correctly modeled this
> deficiency. This analysis would be straightforward to include in the
> current work.

We added a paragraph discussing this issue in the conclusion section (the 2nd paragraph). It is less surprising if we take the perspective that the listener is simulating the speaker by using his own production module, which has the same deficiency.

> Second, in future work, it may be interesting to
> determine if more perspective-taking can be experimentally
> induced, e.g. by having speakers first play as listeners, and
> vice-versa.

   This is indeed an interesting and important empirical question. We did not mention it as this is not the focus 
of the paper.

> In evaluating d), the authors should also consider whether
> listeners model speakers as having the same degree of rationality as
> them (i.e. whether there is only a single lambda parameter in the action-oriented
> listener model).

  We did consider this possibility and did not include it due to space reasons. The issue is now discussed in the 3rd paragraph of the conclusion section.

> It is also not clear whether, in evaluating c), the paper correctly
> models the speaker who believes that the listener is
> action-oriented. The paper uses equation 9) in order to model this speaker. This equation represents
> a speaker who believes that the listener chooses referents in
> proportion to their posterior probability. However, if the speaker is
> modeling the listener as a decision-maker, then it seems more accurate
> for the speaker to represent the listener as a soft-max optimizer, as
> in Equation 14).

  This is a good point and the issue is complicated. We added some discussion in footnote 9. On the conceptual level, it might be argued that choosing a referent according to perceptual salience does not involve rational decision making and thus there may be no need to introduce a soft-max step for the literal listener. On the empirical side, since the perceptual salience was measured using the forced choice paradigm, it is already action-oriented, so the problem is rather about whether we need to introduce a latent salience prior for the belief-oriented models. A very informative next step would be to compare the results for salience conditions from the betting and forced choice paradigms and see whether an extra soft-max step is needed.

Reviewer B:

Reviewer B was unhappy with the presentation of the formalism, because it seemed cluttered and intelligible to him. We reformulated many passages and added in particular footnotes 5 and 6 to discuss the motivation of the exponential choice rule and its relation to the power rule.

Reviewer B wanted to see more details about the experimental design of the original paper. We therefore included this in Section 2 with a few paragraphs, a figure that depicts all situations used and footnotes detailing the original experiment and foreshadowing the changes in our replication.
